---
title: "Seurat2Output_GITHUB_FINAL"
date: "9/21/2022"
author: "HDR"
output: html_document
---

LIBRARIES 
```{r}
library(Seurat)
library(SeuratData)
library(SeuratWrappers)
library(rliger)
library(patchwork)
library(warrenlabSC)
library(ggplot2)
library(cowplot)
library(EnhancedVolcano)
library(ggpubr)
library(dplyr)
library(ComplexHeatmap)
library(gprofiler2)
library(circlize)
library(SingleR)
library(celldex)
library(edgeR)
library(reshape)
library(tidyverse)
library(plyr)
library(RColorBrewer)
library(ggpubr)
library(gridExtra)
library(metap)
library(eulerr)
library(Matrix)
library(DropletUtils)
library(batchelor)
library(miceadds)
require(scater)
require(scran)
require(BiocParallel)
require(BiocNeighbors)
require(data.table)
library(monocle3)
library(enrichplot)
library(DOSE) 
library(CellChat)
library(nichenetr)
```

SEURAT OBJECT CREATION
```{r}
#Load the filtered_feature_bc_matrix output from CellRanger AGGR
data <- Read10X(data.dir = "path")
#Initialize the Seurat object with the raw (non-normalized) data
object <- CreateSeuratObject(counts = data, project= "integration", min.cells = 2, min.features = 200)
#add percent_mt to metadata as quality metric 
object[["percent_mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")
#add meta.data from the aggregation.csv output from CellRanger AGGR
meta.data <- read.csv("path")
object <- add.meta.data(object, meta.data)
#remove molecule_h5 column to reduce matrix size if desired 
object@meta.data$molecule_h5 <- NULL
#visualize quality metrics
VlnPlot(object, features = "nFeature_RNA", group.by = "sample_id", pt.size = 0)+theme(axis.text.x = element_text(size = 6))
VlnPlot(object, features = "percent_mt", group.by = "sample_id", pt.size = 0)+theme(axis.text.x = element_text(size = 6))
#subset seurat object based on quality metrics 
object <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 3000 & percent_mt < 12)
table(object@meta.data$sample_id)
#integrate 
# split the dataset into a list of seurat objects by 10X well 
list <- SplitObject(object, split.by = "sample_id")
# normalize and identify variable features for each dataset independently
list <- lapply(X = list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = list)
#find anchors 
anchors <- FindIntegrationAnchors(object.list = list, anchor.features = features)
#this command creates an 'integrated' data assay
combined <- IntegrateData(anchorset = anchors)
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(combined) <- "integrated"
# Run the standard workflow for visualization and clustering
combined <- ScaleData(combined, verbose = FALSE)
combined <- RunPCA(combined, npcs = 30, verbose = FALSE)
combined <- RunUMAP(combined, reduction = "pca", dims = 1:30)
combined <- FindNeighbors(combined, reduction = "pca", dims = 1:30)
combined <- FindClusters(combined, resolution = 0.5)
#scale RNA assay
DefaultAssay(combined) <- "RNA"
combined <- ScaleData(combined, assay = "RNA")
#Save seurat object 
save(combined,file="combined.RData")
```

LOAD DATA 1
```{r}
#load seurat object
load(file = "combined.RData")
```

MANUAL CELLTYPING
```{r}
DefaultAssay(combined) <- "RNA"
#PROG
PROG_Markers <- c('NES','PAX6','SOX2',"SOX1")
pdf(paste0("PROG_feature.pdf"))
FeaturePlot(combined, features = PROG_Markers, combine = TRUE, order = F, raster = F) +theme(text = element_text(face = "bold"))
dev.off()
combined <- AddModuleScore(object=combined, features = PROG_Markers, ctrl = 5, name = "PROGfeatures")
pdf(paste0("module_PROG_feature.pdf"))
FeaturePlot(combined, features = "PROGfeatures1", min.cutoff = 0, order = T, raster = F)+ labs(title = "PROG Markers")+theme(text = element_text(face = "bold"))
dev.off()
#CYC
CYC_Markers <- c("CENPF","TOP2A","HMGB2","UBE2C")
pdf(paste0("CYC_feature.pdf"))
FeaturePlot(combined, features = CYC_Markers, combine = TRUE, order = F, raster = F)+theme(text = element_text(face = "bold"))
dev.off()
combined <- AddModuleScore(object=combined, features = CYC_Markers, ctrl = 5, name = "CYCfeatures")
pdf(paste0("module_CYC_feature.pdf"))
FeaturePlot(combined, features = "CYCfeatures1", min.cutoff = 0, order = T, raster = F)+ labs(title = "CYC Markers")+theme(text = element_text(face = "bold"))
dev.off()
#NEU
NEU_Markers <- c("GAP43",'SNAP25','MAPT','STMN2')
pdf(paste0("NEU_feature.pdf"))
FeaturePlot(combined, features = NEU_Markers, combine = TRUE, order = F, raster = F)+theme(text = element_text(face = "bold"))
dev.off()
combined <- AddModuleScore(object=combined, features = NEU_Markers, ctrl = 5, name = "NEUfeatures")
pdf(paste0("module_NEU_feature.pdf"))
FeaturePlot(combined, features = "NEUfeatures1", min.cutoff = 0, order = T, raster = F) + labs(title = "PAN_NEU Markers")+theme(text = element_text(face = "bold"))
dev.off()
#EXCITATORY 
GLU_Markers <- c("NEFM","ONECUT2", "GRID2","GRIA4","SLC17A6","GRIN2A")
pdf(paste0("GLU_feature.pdf"))
FeaturePlot(combined, features = GLU_Markers, combine = TRUE, order = T, raster = F)+theme(text = element_text(face = "bold"))
dev.off()
combined <- AddModuleScore(object=combined, features = GLU_Markers, ctrl = 5, name = "GLUfeatures")
pdf(paste0("module_GLU_feature.pdf"))
FeaturePlot(combined, features = "GLUfeatures1", min.cutoff = 0, order = T, raster = F)+ labs(title = "EX_NEU Markers")+theme(text = element_text(face = "bold"))
dev.off()
#INHIBITORY 
GABA_Markers <- c("GAD1","GAD2","SLC32A1","CNR1")
pdf(paste0("GABA_feature.pdf"))
FeaturePlot(combined, features = GABA_Markers, combine = TRUE, order = T, raster = F)+theme(text = element_text(face = "bold"))
dev.off()
combined<- AddModuleScore(object=combined, features = GABA_Markers, ctrl = 5, name = "GABAfeatures")
pdf(paste0("module_GABA_feature.pdf"))
FeaturePlot(combined, features = "GABAfeatures1", min.cutoff = 0, order = T, raster = F) + labs(title = "IN_NEU Markers")+theme(text = element_text(face = "bold"))
dev.off()
#ASC
ASC_Markers <- c("SLC1A3","VIM","SPARC","PEA15","VCAM1","CLU")
pdf(paste0("ASC_feature.pdf"))
FeaturePlot(combined, features = ASC_Markers, combine = TRUE, order = F, raster = F)+theme(text = element_text(face = "bold"))
dev.off()
combined <- AddModuleScore(object=combined, features = ASC_Markers, ctrl = 5, name = "ASCfeatures")
pdf(paste0("module_ASC_feature.pdf"))
FeaturePlot(combined, features = "ASCfeatures1", min.cutoff = 0, order = T, raster = F)+ labs(title = "ASC Markers")+theme(text = element_text(face = "bold"))
dev.off()
```

AUTOMATED CELLTYPING
```{r}
#Get data from seurat object
counts <- GetAssayData(combined)
#For human data, call the human primary cell altas as reference from celldex, can use blueprint for Blueprint Epigenomics and Encode data set.
#Can use immgen and mouse.rnaseq for mouse data as well. Check documentation.
hpca.se <- HumanPrimaryCellAtlasData()
#can check whats in the reference by 
#hpca.se$label.main
#hpca.se$label.fine
#Set default assay
DefaultAssay(combined)
#Use the human cell altas as reference to assign main label cell (broad cell type) type to Seurat object
#pbmc_hpca <- SingleR(test=counts, ref= hpca.se, labels = hpca.se$label.main,assay.type.test="intergrated")
#Use the human cell altas as reference to assign fine label cell (fine cell type) type to Seurat object
pbmc_hpca <- SingleR(test=counts, ref= hpca.se, labels = hpca.se$label.fine,assay.type.test="intergrated")
combined <- AddMetaData(object = combined, metadata = pbmc_hpca$first.labels , col.name = "SingleR_Celltype")
#check what has been assigned to each cell in table format
table(combined@meta.data$seurat_clusters,combined@meta.data$SingleR_Celltype)
#plot
combined.active <- SetIdent(combined, value= "SingleR_Celltype")
pdf(paste0("SingleR_Celltype_fine.pdf"), width = 20, height = 20)
DimPlot(combined.active,  shuffle = T, raster = F)
dev.off()
```

CELLTYPE DOCUMENTATION 
```{r}
#celltype assignment 
Idents(combined)<- combined$seurat_clusters 
combined <- RenameIdents(combined, `0` = "IN_NEU", `1` = "ASC", `2` = "IN_NEU", `3` = "EX_NEU", `4` = "NEU_X", `5` = "ASC", `6` = "EX_NEU", `7` = "IN_NEU", `8` = "NEU_X", `9` = "EX_NEU",  `10` = "NEU_P", `11` = "EX_NEU", `12` = "ASC", `13` = "IN_NEU", `14` = "NEU_P", `15` = "NEU_P", `16` = "ASC_P",`17` = "ASC_P", `18` = "ASC", `19` = "NEU_P",`20` = "IN_NEU",`21` = "ASC_P")
combined$celltype <- Idents(combined)
#new metadata columns 
combined$celltype.sample <- paste(combined$celltype, combined$sample_id, sep = "_")
combined$celltype.condition <- paste(combined$celltype, combined$condition, sep = "_")
combined$celltype.batch <- paste(combined$celltype, combined$batch, sep = "_")
combined$celltype.week <- paste(combined$celltype, combined$week, sep = "_")
combined$celltype.treat <- paste(combined$celltype, combined$treatment, sep = "_")
combined$condition.treat <- paste(combined$condition, combined$treatment, sep = "_")
combined$condition.treat.week <- paste(combined$condition, combined$treatment,combined$week, sep = "_")
combined$celltype.batch.condition <- paste(combined$celltype, combined$batch, combined$condition, sep = "_")
combined$celltype.condition.week.treat <- paste(combined$celltype, combined$condition,combined$week,combined$treatment,sep = "_")

#Save named seurat object
save(combined,file="combined_named.RData")
```

LOAD DATA2
```{r}
#load seurat object
load(file = "combined_named.RData")

##DEFINE COLORS AND PALETTES##
#celltype colors 
col_EX_NEU <-  "aquamarine4"  
col_IN_NEU <- "aquamarine3" 
col_NEU_P <- "aquamarine1"
col_NEU_X <-"azure1"
col_ASC <-"tomato4"
col_ASC_P <-"tomato2" 
col_ASC_X <-"azure2" 
col_CYC <-"gold1"
#conditions colors
col_ctr_v_w1<- "#DEEBF7"
col_ctr_v_w2<- "#6BAED6"
col_ctr_v_w3<- "#084594"
col_otau_v_w1<-  "#FDD0A2"
col_otau_v_w2<- "#FD8D3C"
col_otau_v_w3<- "#D94801"
col_asyn_v_w3<- "#FCBBA1" 
col_ctr_pu_w2<- "#74C476"
col_ctr_pu_w3<- "#238B45"
col_otau_pu_w2<- "#9E9AC8"
col_otau_pu_w3<- "#6A51A3"
#color  palettes 
cellcolors <- c(col_ASC_P,col_ASC,col_NEU_P,col_EX_NEU,col_IN_NEU,col_NEU_X)
label_colors_ctr <- c(col_ctr_v_w1,col_ctr_v_w2,col_ctr_v_w3)
condition_treat_week_colors <- c(col_ctr_v_w1,col_otau_v_w1,col_ctr_v_w2,col_otau_v_w2,col_ctr_v_w3,col_otau_v_w3,col_ctr_pu_w2,col_otau_pu_w2,col_ctr_pu_w3,col_otau_pu_w3,col_asyn_v_w3)
condition_treat_week_colors_otau_only <- c(col_ctr_v_w1,col_otau_v_w1,col_ctr_v_w2,col_otau_v_w2,col_ctr_v_w3,col_otau_v_w3,col_ctr_pu_w2,col_otau_pu_w2,col_ctr_pu_w3,col_otau_pu_w3)
condition_treat_week_colors_otau_vehicle_only <- c(col_ctr_v_w1,col_otau_v_w1,col_ctr_v_w2,col_otau_v_w2,col_ctr_v_w3,col_otau_v_w3)
condition_treat_week_colors_ctr_vehicle_only <- c(col_ctr_v_w1,col_ctr_v_w2,col_ctr_v_w3)
#label palettes
cond_labels_all <- c("7 DIV3D Control Vehicle","7 DIV3D oTAU Vehicle","14 DIV3D Control Vehicle","14 DIV3D oTAU Vehicle","21 DIV3D Control Vehicle","21 DIV3D oTAU Vehicle","14 DIV3D Control PU","14 DIV3D oTAU PU","21 DIV3D Control PU","21 DIV3D oTAU PU","21 DIV3D aSYN-PFF Vehicle")
cond_labels_otau_only <- c("7 DIV3D Control Vehicle","7 DIV3D oTAU Vehicle","14 DIV3D Control Vehicle","14 DIV3D oTAU Vehicle","21 DIV3D Control Vehicle","21 DIV3D oTAU Vehicle","14 DIV3D Control PU","14 DIV3D oTAU PU","21 DIV3D Control PU","21 DIV3D oTAU PU")
cond_labels_otau_vehicle_only <- c("7 DIV3D Control Vehicle","7 DIV3D oTAU Vehicle","14 DIV3D Control Vehicle","14 DIV3D oTAU Vehicle","21 DIV3D Control Vehicle","21 DIV3D oTAU Vehicle")
cond_labels_ctr_vehicle_only <- c("7 DIV3D Control Vehicle","14 DIV3D Control Vehicle","21 DIV3D Control Vehicle")
sample_id_labels <-  c("Rep1 7 DIV3D Control Vehicle","Rep1 7 DIV3D oTAU Vehicle","Rep1 14 DIV3D Control Vehicle","Rep1 14 DIV3D oTAU Vehicle","Rep1 21 DIV3D Control Vehicle","Rep1 21 DIV3D oTAU Vehicle","Rep1 14 DIV3D Control PU","Rep1 14 DIV3D oTAU PU","Rep1 21 DIV3D Control PU","Rep1 21 DIV3D oTAU PU","Rep2 7 DIV3D Control Vehicle","Rep2 7 DIV3D oTAU Vehicle","Rep2 14 DIV3D Control Vehicle","Rep2 14 DIV3D oTAU Vehicle","Rep3 21 DIV3D Control Vehicle","Rep3 21 DIV3D oTAU Vehicle","Rep4 21 DIV3D Control Vehicle","Rep4 21 DIV3D oTAU Vehicle","Rep4 21 DIV3D PFFaSYN Vehicle","Rep5 21 DIV3D Control Vehicle","Rep5 21 DIV3D oTAU Vehicle")
#cell subset palettes
condition_treat_week_cells_otau_only <- c("ctr_v_w1","otau_v_w1","ctr_v_w2","otau_v_w2","ctr_v_w3","otau_v_w3","ctr_pu_w2","otau_pu_w2","ctr_pu_w3","otau_pu_w3")
condition_treat_week_cells_otau_vehicle_only <- c("ctr_v_w1","otau_v_w1","ctr_v_w2","otau_v_w2","ctr_v_w3","otau_v_w3")
condition_treat_week_cells_ctr_vehicle_only <- c("ctr_v_w1","ctr_v_w2","ctr_v_w3")

##RELEVEL##
#celltype
Idents(combined) <- "celltype"
# Define an order of cluster identities
cell_levels <- c("ASC_P","ASC","NEU_P","EX_NEU","IN_NEU","NEU_X")
# Relevel object@ident
combined@meta.data$celltype <- factor(x = combined@meta.data$celltype, levels = cell_levels)

#condition.treat.week
Idents(combined) <- "condition.treat.week"
# Define an order of cluster identities
condition_treat_week_levels <- c("ctr_v_w1","otau_v_w1","ctr_v_w2","otau_v_w2","ctr_v_w3","otau_v_w3","ctr_pu_w2","otau_pu_w2","ctr_pu_w3","otau_pu_w3","asyn_v_w3")
# Relevel object@ident
combined@meta.data$condition.treat.week <- factor(x = combined@meta.data$condition.treat.week, levels = condition_treat_week_levels)

#sample_id
Idents(combined) <- "sample_id"
# Define an order of cluster identities
sample_id_levels <-  c("b1_w1_ctr_v","b1_w1_otau_v","b1_w2_ctr_v","b1_w2_otau_v","b1_w3_ctr_v","b1_w3_otau_v","b1_w2_ctr_pu","b1_w2_otau_pu","b1_w3_ctr_pu","b1_w3_otau_pu","b2_w1_ctr_v", "b2_w1_otau_v","b2_w2_ctr_v","b2_w2_otau_v","b3_w3_ctr_v","b3_w3_otau_v","b4_ctr_v","b4_otau_v","b4_asyn_v", "b5_w3_ctr_v", "b5_w3_otau_v")
# Relevel object@ident
combined@meta.data$sample_id <- factor(x = combined@meta.data$sample_id, levels = sample_id_levels)

#week
Idents(combined) <- "week"
# Define an order of cluster identities
week_levels <- c("w1","w2","w3")
# Relevel object@ident
combined@meta.data$week <- factor(x = combined@meta.data$week, levels = week_levels)

```

CELLTYPE HEATMAP
```{r}
combined <- SetIdent(combined, value ="celltype")
markers <- FindAllMarkers(combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
saveRDS(markers, paste0("celltype_markers_2.rds"))
write.csv(markers, file = "celltype_markers_2.csv")
markers <- read.delim2(file = "celltype_markers_2.csv", sep = ",")
top10 <- group_by(markers,cluster)
top10 <- top_n(top10, n = 20, wt = avg_log2FC)
plot <- DoHeatmap(subset(combined, downsample=1000), features = top10$gene, size = 0, angle = 0, hjust = 0.5, label = TRUE, group.by = "celltype", assay = "RNA", slot = "scale.data", group.colors = cellcolors,draw.lines = T) + theme(axis.text.y = element_text(size = 2))
pdf("celltype_heatmap_2.pdf")
print(plot)
dev.off()
```

UMAP VISUALIZATIONS
```{r}
#set integrated default assay 
DefaultAssay(combined) <- "integrated"
#seurat clusters 
pdf("cluster_umap.pdf")
DimPlot(combined, reduction = "umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE, label.size = 8, raster = F)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15) )
dev.off()
#celltypes 
pdf("celltype_umap.pdf")
DimPlot(combined, label = TRUE, cols = cellcolors, group.by = "celltype", label.size = 8, raster = F)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15))
dev.off()
#conditions 
#all conditions 
pdf("condition_umap_all.pdf")
DimPlot(combined, reduction = "umap", group.by = "condition.treat.week", shuffle = TRUE, raster=F)+scale_color_manual(labels= cond_labels_all,values= condition_treat_week_colors)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15))
dev.off()
#otau only  
Idents(combined) <- "condition.treat.week"
sub_combined <- subset(combined,idents=condition_treat_week_cells_otau_only)
pdf("condition_umap_otau_only.pdf")
DimPlot(sub_combined, reduction = "umap",group.by = "condition.treat.week",shuffle = TRUE, raster=F)+scale_color_manual(labels= cond_labels_otau_only,values= condition_treat_week_colors_otau_only)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15))
dev.off()
#otau vehicle only  
Idents(combined) <- "condition.treat.week"
sub_combined <- subset(combined,idents=condition_treat_week_cells_otau_vehicle_only)
pdf("condition_umap_otau_vehicle_only.pdf")
DimPlot(sub_combined, reduction = "umap",group.by = "condition.treat.week",shuffle = TRUE, raster=F)+scale_color_manual(labels= cond_labels_otau_vehicle_only,values= condition_treat_week_colors_otau_vehicle_only)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15))
dev.off()
#ctr vehicle only  
Idents(combined) <- "condition.treat.week"
sub_combined <- subset(combined,idents=condition_treat_week_cells_ctr_vehicle_only)
pdf("condition_umap_ctr_vehicle_only.pdf")
DimPlot(sub_combined, reduction = "umap",group.by = "condition.treat.week",shuffle = TRUE, raster=F)+scale_color_manual(labels= cond_labels_ctr_vehicle_only,values= condition_treat_week_colors_ctr_vehicle_only)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15))
dev.off()
#maturation gradient 
Idents(combined) <- "week"
pdf("maturation_gradient_umap.pdf")
DimPlot(combined, label = F, shuffle = T, raster = F)+theme(text = element_text(face = "bold"),title = element_text(size = 0), axis.text = element_text(size = 15)) +
  scale_color_manual(labels = c("DIV3D 7", "DIV3D 14", "DIV3D 21"),  values = c("yellow","red","purple") )
dev.off()
#naturation legend
col_fun = colorRamp2(c(7,14,21), c("yellow","red","purple"))
lgd = Legend(col_fun = col_fun, title = "Maturity", at = c(7,14,21),labels = c("low", "median", "high"), direction = "horizontal")
pdf("legend_maturity.pdf")
draw(lgd)
dev.off()
#Feature Plots 
DefaultAssay(combined) <- "RNA"
genes <- c("VIM","SLC1A3","SOX2","MAP2","NES","TOP2A","GAD1","GRIA4","SYP","SYN1","DLG4","VAMP2","SNAP25","nFeature_RNA","nCount_RNA","percent_mt")
for(gn in genes){
plot<- FeaturePlot(combined, features=gn, order = T, raster = F)+theme(text = element_text(face = "bold"))
pdf(paste0(gn,"_featureplot.pdf"))
print(plot)
dev.off()
}
#VLN Plots
p <- VlnPlot(combined, features = "nFeature_RNA", group.by = "sample_id", pt.size = 0, y.max = 5000)+ NoLegend()+scale_x_discrete(labels=sample_id_labels)+theme(text = element_text(face = "bold"),axis.title.x = element_text(size = 0), axis.text = element_text(size = 10, face = "bold"), axis.text.x.bottom = element_text(angle = 90)) 
pdf("nFeature_vln.pdf")
plot(p)
dev.off()
p <- VlnPlot(combined, features = "nCount_RNA", group.by = "sample_id", pt.size = 0, y.max =15000)+ NoLegend()+scale_x_discrete(labels=sample_id_labels)+theme(text = element_text(face = "bold"),axis.title.x = element_text(size = 0), axis.text = element_text(size = 10, face = "bold"), axis.text.x.bottom = element_text(angle = 90)) 
pdf("nCount_vln.pdf")
plot(p)
dev.off()
p <- VlnPlot(combined, features = "percent_mt", group.by = "sample_id", pt.size = 0, y.max =20)+ NoLegend()+scale_x_discrete(labels=sample_id_labels)+theme(text = element_text(face = "bold"),axis.title.x = element_text(size = 0), axis.text = element_text(size = 10, face = "bold"), axis.text.x.bottom = element_text(angle = 90)) 
pdf("percent_mt_vln.pdf")
plot(p)
dev.off()
```

CELLTYPE BARPLOT
```{r}
#ALL
#dataframe setup
combined <- SetIdent(combined, value ="celltype")
ident <- c("ctr_v_w1","otau_v_w1","ctr_v_w2","otau_v_w2","ctr_v_w3","otau_v_w3","ctr_pu_w2","otau_pu_w2","ctr_pu_w3","otau_pu_w3")
ns<-length(ident)
clusters<-levels(combined)
ncl<-length(clusters)
counts<-matrix(0,nrow=ncl,ncol=ns)
rownames(counts)<-clusters
colnames(counts)<-ident
#count per celltype
for (s in ident) {
  for (cl in clusters) {
    counts[cl,s]<-sum(combined@meta.data$condition.treat.week==s & combined@meta.data$celltype==cl)
  }
}
barplot_label_colors <- c(col_otau_pu_w3,col_ctr_pu_w3,col_otau_pu_w2,col_ctr_pu_w2,col_otau_v_w3,col_ctr_v_w3,col_otau_v_w2,col_ctr_v_w2,col_otau_v_w1,col_ctr_v_w1)
colnames(counts) <- c("7     -     -","7     +     -","14     -     -","14     +     -","21     -     -","21     +     -","14     -     +","14     +     +","21     -     +","21     +     +")
counts <- counts[,order(ncol(counts):1)]
counts <- prop.table(counts,margin = 2)
counts <- counts*100
counts <- as.data.frame(counts)
counts$ID <- rownames(counts)
counts.melt <- melt(counts, id.var = 'ID')
counts.melt <- within(counts.melt, ID <- factor(ID,
    cell_levels, ordered = TRUE))
labelcolors <- "black"
barfill <- cellcolors
plot<- ggplot(counts.melt, aes(y = variable, x = value, fill = ID,)) +
    geom_bar(stat = 'identity', width = 0.5,colour="black") +
  scale_fill_manual(values = barfill )+
    theme(text = element_text(face = "bold"),aspect.ratio = .8,legend.title=element_blank(),
          legend.text = element_text(size=18),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(), axis.text.x = element_blank(),
          axis.text.y = element_text(color = labelcolors,face = "bold",size = 15),
          axis.ticks = element_blank(),panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())
pdf("all_cellcount_barplot.pdf", width = 5, height = 2.5)
print(plot)
dev.off()

#VEHICLE ONLY
#dataframe setup
combined <- SetIdent(combined, value ="celltype")
ident <- c("ctr_v_w1","otau_v_w1","ctr_v_w2","otau_v_w2","ctr_v_w3","otau_v_w3")
ns<-length(ident)
clusters<-levels(combined)
ncl<-length(clusters)
counts<-matrix(0,nrow=ncl,ncol=ns)
rownames(counts)<-clusters
colnames(counts)<-ident
#count per celltype
for (s in ident) {
  for (cl in clusters) {
    counts[cl,s]<-sum(combined@meta.data$condition.treat.week==s & combined@meta.data$celltype==cl)
  }
}
barplot_label_colors <- c(col_otau_v_w3,col_ctr_v_w3,col_otau_v_w2,col_ctr_v_w2,col_otau_v_w1,col_ctr_v_w1)
colnames(counts) <- c("7     -","7     +","14     -","14     +","21     -","21     +")
counts <- counts[,order(ncol(counts):1)]
counts <- prop.table(counts,margin = 2)
counts <- counts*100
counts <- as.data.frame(counts)
counts$ID <- rownames(counts)
counts.melt <- melt(counts, id.var = 'ID')
counts.melt <- within(counts.melt, ID <- factor(ID,
    cell_levels, ordered = TRUE))
labelcolors <- "black"
barfill <- cellcolors
plot<- ggplot(counts.melt, aes(y = variable, x = value, fill = ID,)) +
    geom_bar(stat = 'identity', width = 0.5,colour="black") +
  scale_fill_manual(values = barfill )+
    theme(text = element_text(face = "bold"),aspect.ratio = .8,legend.title=element_blank(),
          legend.text = element_text(size=10),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(), axis.text.x = element_blank(),
          axis.text.y = element_text(color = labelcolors,face = "bold",size = 10),
          axis.ticks = element_blank(),panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())
pdf("vehicle_cellcount_barplot.pdf", width = 5, height = 2.5)
print(plot)
dev.off()

#CONTROL
#dataframe setup
combined <- SetIdent(combined, value ="celltype")
ident <- c("ctr_v_w1","ctr_v_w2","ctr_v_w3")
ns<-length(ident)
clusters<-levels(combined)
ncl<-length(clusters)
counts<-matrix(0,nrow=ncl,ncol=ns)
rownames(counts)<-clusters
colnames(counts)<-ident
#count per celltype
for (s in ident) {
  for (cl in clusters) {
    counts[cl,s]<-sum(combined@meta.data$condition.treat.week==s & combined@meta.data$celltype==cl)
  }
}
colnames(counts) <- c("7", "14", "21")
barplot_label_colors_ctr <- c(col_ctr_v_w3,col_ctr_v_w2,col_ctr_v_w1)
counts <- prop.table(counts,margin = 2)
counts <- counts*100
counts <- as.data.frame(counts)
counts$ID <- rownames(counts)
counts.melt <- melt(counts, id.var = 'ID')
counts.melt <- within(counts.melt, ID <- factor(ID,
    cell_levels, ordered = TRUE))
labelcolors <- "black"
barfill <- cellcolors

plot <- ggplot(counts.melt, aes(y = variable, x = value, fill = ID,)) +
    geom_bar(stat = 'identity', width = 0.5,colour="black") +
  scale_fill_manual(values = barfill )+
    theme(legend.position = "right", text = element_text(face = "bold"),aspect.ratio = 1,legend.title=element_blank(),
          legend.text = element_text(size=10),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(), 
          axis.text.y = element_blank(),
          axis.text.x = element_text(color = labelcolors,face = "bold",size = 0),
          axis.ticks = element_blank(),panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())
plot2 <- plot + coord_flip()
pdf("ctr_cellcount_barplot.pdf", width = 5, height = 5)
print(plot2)
dev.off()


#colored by batch replicate
Idents(combined) <- "batch"
#table data 
t <- table(combined@meta.data$batch, combined@meta.data$celltype)
#convert to percentage of total
t <- prop.table(t,margin = 1)
t <- t*100
rownames(t) <- c("C","D","E","A","B")
t <- t[order(rownames(t)), ]
#document
#write_csv(t, "celltype_counts.csv")
# melt the data frame for plotting
data.m <- melt(t, id.vars='celltype')
colnames(data.m) <- c("Replicate","Cell_Type","Percent_Composition")
#plot
plot<- ggplot(data.m, aes(y = Replicate, x = Percent_Composition, fill = Cell_Type)) +geom_bar(stat = 'identity', width = 1,colour="black") + coord_flip()+theme(legend.text = element_text(size = 15, colour = "black", face = "bold"), legend.title = element_text(size=15, face = "bold"),axis.text = element_text(size = 15, colour = "black", angle = 20, face = "bold"), axis.title.y = element_text(size = 15, face = "bold"), axis.title.x = element_text(size =15, face = "bold"),legend.position = "bottom", legend.direction = "horizontal")+ theme(panel.background = element_blank())

pdf(paste0("batch_celltype_counts.pdf"))
print(plot)
dev.off()
```

NEURON LOSS BARPLOT
```{r}
#subset to otau w1,w2,w3
Idents(combined)<-"condition.treat.week"
subset <- subset(combined, idents=c("otau_v_w1","otau_v_w2","otau_v_w3","ctr_v_w1","ctr_v_w2","ctr_v_w3"))

#subset
subset$batch.condition.week.treat <- paste(subset$batch_id, subset$condition,subset$week,subset$treatment,sep = "_")
Idents(subset) <- "batch.condition.treat.week"
#table data
t <- table(subset@meta.data$batch.condition.week.treat, subset@meta.data$celltype)
#convert to percentage of total
t <- prop.table(t,margin = 2)
t <- t*100

t <- t[,c("EX_NEU")]

data.m <- melt(t, id.vars=rownames)
data.m <- cbind(rownames(data.m), data.frame(data.m, row.names=NULL))
colnames(data.m) <- c("Condition","Percent")

data.m$Week <- c("7 DIV3D", "14 DIV3D", "21 DIV3D", "7 DIV3D", "14 DIV3D", "21 DIV3D", "7 DIV3D", "14 DIV3D", "7 DIV3D", "14 DIV3D", "21 DIV3D", "21 DIV3D", "21 DIV3D", "21 DIV3D", "21 DIV3D", "21 DIV3D")
data.m$Treatment <- c("Control", "Control", "Control","AstTau", "AstTau", "AstTau","Control", "Control","AstTau", "AstTau", "Control","AstTau", "Control","AstTau", "Control","AstTau")

data.m$Treatment <- factor(data.m$Treatment, levels = c("Control","AstTau"))
data.m$Week <- factor(data.m$Week, levels = c("7 DIV3D", "14 DIV3D", "21 DIV3D"))

#plot
plot<- ggplot(data.m, aes( y=Percent, x=Week, fill = Treatment)) +geom_bar(stat = 'identity', width = 0.75,colour="black",position=position_dodge(.8)) +theme(legend.text = element_text(size = 15, colour = "black", face = "bold"), legend.title = element_text(size=15, face = "bold"),axis.text = element_text(size = 15, colour = "black", angle = 0, face = "bold"), axis.title.y = element_text(size = 15, face = "bold"), axis.title.x = element_text(size =15, face = "bold"),legend.position = "bottom", legend.direction = "horizontal")+ theme(panel.background = element_blank())+
  xlab("Week") + ylab("Percent EX_NEU")+
  scale_fill_manual(values=c(col_ctr_v_w3,col_otau_v_w3)) 

pdf(paste0("glu_celltype_counts.pdf"), width = 5)
print(plot)
dev.off()

```

CELLTYPE DOTPLOT
```{r}
diff_groups <- c("CYC","ASC","PAN-NEU","GABA","GLU","SYNAPSE")
features <-c("PAX6","NES","SOX2","TOP2A","VIM","SLC1A3","PEA15","SPARC","GAP43","STMN2","MAPT","SNAP25","GAD1","GAD2","SLC32A1","SLC38A1","GRIA4","SLC17A6","SYP","SYT1", "DLG4")
plot<- DotPlot(combined, features = features, cluster.idents = F, cols = c("blue", "red"), scale.by = "size")+
  theme(text=element_text(face="bold"),axis.text.x = element_text(size=10, angle = 90), legend.position = "bottom", legend.text = element_text(size = 10), legend.title = element_text(size = 10))+
  geom_rect(aes(xmin=0,xmax=4.5,ymin=6.5,ymax=7.5),fill=col_CYC,alpha=0.1,color="black")+
  geom_rect(aes(xmin=4.5,xmax=8.5,ymin=6.5,ymax=7.5),fill=col_ASC,alpha=0.1,color="black")+
  geom_rect(aes(xmin=8.5,xmax=12.5,ymin=6.5,ymax=7.5),fill=col_NEU_P,alpha=0.1,color="black")+
  geom_rect(aes(xmin=12.5,xmax=15.5,ymin=6.5,ymax=7.5),fill=col_IN_NEU,alpha=0.1,color="black")+
  geom_rect(aes(xmin=15.5,xmax=18.5,ymin=6.5,ymax=7.5),fill=col_EX_NEU,alpha=0.1,color="black")+
  geom_rect(aes(xmin=18.5,xmax=21.5,ymin=6.5,ymax=7.5),fill=col_NEU_X,alpha=0.1,color="black")+
  annotate(geom = "text", x = c(2.5,6.5,10.5,14,17,20), y =7, label = diff_groups, size = 4)+
  labs(y="Celltype",x="Genes")
pdf("diff_dotplot.pdf")
print(plot)
dev.off()
```

CELLTYPE STACKED VLNPLOT
```{r}
#NEW FUNCTION

## remove the x-axis text and tick
## plot.margin to adjust the white space between each plot.
## ... pass any arguments to VlnPlot in Seurat
modify_vlnplot<- function(obj, 
                          feature, 
                          pt.size = 0.0,
                          plot.margin = unit(c(0, 0, 0, 0), "cm"),
                          ...) {
  p<- VlnPlot(obj, features = feature, pt.size = pt.size, ... )  + 
    xlab("") + ylab(feature) + ggtitle("") + 
    theme(legend.position = "none", 
          plot.title = element_text(size = 0),plot.caption = element_text(size = 0),
          axis.text.x = element_text(size=0), axis.title.x = element_text(size=0),
          axis.ticks.length.x = unit(0, "pt"),
          axis.ticks.x = element_blank(),
          axis.title.y = element_text(size = rel(1), angle = 0, face = "bold"), 
          axis.text.y = element_text(size = rel(1), face = "bold"),
          plot.margin = plot.margin ) 
  return(p)
}

## extract the max value of the y axis
extract_max<- function(p){
  ymax<- max(ggplot_build(p)$layout$panel_scales_y[[1]]$range$range)
  return(ceiling(ymax))
}

## main function
StackedVlnPlot<- function(obj, features,
                          pt.size = 0.0, 
                          plot.margin = unit(c(0, 0, 0, 0), "cm"),
                          ...) {
  
  plot_list<- purrr::map(features, function(x) modify_vlnplot(obj = obj,feature = x, ...))
  
  # Add back x-axis title to bottom plot. patchwork is going to support this?
  plot_list[[length(plot_list)]]<- plot_list[[length(plot_list)]] +
    theme(axis.text.x=element_text(size = 15, angle = 35, face= "bold"), axis.ticks.x = element_line())
  
  # change the y-axis tick to only max value 
  ymaxs<- purrr::map_dbl(plot_list, extract_max)
  plot_list<- purrr::map2(plot_list, ymaxs, function(x,y) x +
                            scale_y_continuous(breaks = c(y)) +
                            expand_limits(y = y))

  p<- patchwork::wrap_plots(plotlist = plot_list, ncol = 1) 
        
  return(p)
}


#PLOT

features <-c("NES","TOP2A","VIM","SLC1A3","PEA15","GAP43","STMN2","NEFM","SLC38A1","SYP")


combined <- SetIdent(combined, value ="celltype")

plot <- StackedVlnPlot(obj = combined, features = features)


#pdf("cell_stckvln.pdf" )
pdf("cell_stckvln_celltype.pdf", width= 5, height =10 )
print(plot)
dev.off()

```

DIFFERENTIAL GENE EXPRESSION (DEG)
```{r}
celltypes <- c("EX_NEU","IN_NEU","ASC")
conditions <- c("otau","asyn")
weeks <- c("w1","w2","w3")
treatment <- c("v","pu")
for(ct in celltypes){
  for (cond in conditions){
  for(wk in weeks){
   for(trt in treatment){
      tryCatch({
Idents(combined) <- "celltype.condition.week.treat"
response <- FindMarkers(combined, ident.1 = paste0(ct,"_",cond,"_",wk,"_",trt), ident.2 = paste0(ct,"_ctr_",wk,"_",trt), verbose = FALSE, pseudocount.use = 0.0001, assay = "RNA", logfc.threshold = 0.25)
response$p_val_adj_fdr=p.adjust(response$p_val,method="fdr")
#csv
response$name=rownames(response)
response=response[order(response$avg_log2FC),]
write.csv(response, paste0(ct,"_",cond,"_",wk,"_",trt,"_DEG.csv"))
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}

```

SIG DEG CSV
```{r}
celltypes <- c("EX_NEU","IN_NEU","ASC")
conditions <- c("otau","asyn")
weeks <- c("w1","w2","w3")
treatment <- c("v","pu")
#UP
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
     tryCatch({
x <- read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_DEG.csv"))
x <- x[with(x, order(p_val)), ]
x <- x[x$p_val_adj_fdr<0.05,]
x <- x[x$avg_log2FC > 0.25,]
x <- subset(x, select = name)
write_csv(x,paste0(ct,"_",cond,"_",wk,"_",trt,"_SIG_UP_DEG.csv"))
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}
#DOWN
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
     tryCatch({
x <- read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_DEG.csv"))
x <- x[with(x, order(p_val)), ]
x <- x[x$p_val_adj_fdr<0.05,]
x <- x[x$avg_log2FC < -0.25,]
x <- subset(x, select = name)
write_csv(x,paste0(ct,"_",cond,"_",wk,"_",trt,"_SIG_DN_DEG.csv"))
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}
```

DEG TABLE
```{r}
celltypes <- c("EX_NEU","IN_NEU","ASC")
conditions <- c("otau","asyn")
weeks <- c("w1","w2","w3")
treatment <- c("v","pu")
#UP
#set up data frame 
df_UP=data.frame(row.names="genes")
num=numeric()
#set up loops 
for(ct in celltypes){
  for(cond in conditions){
    for(wk in weeks){
      for(treat in treatment){
        tryCatch({
#read file
x_UP <- read.csv(paste0(ct,"_",cond,"_",wk,"_",treat,"_SIG_UP_DEG.csv"))
#count genes/rows
num <- nrow(x_UP)
#bind to data frame 
df_UP=cbind(df_UP,num)
#set colnames 
colnames(df_UP)[ncol(df_UP)]=paste0(ct,"_",cond,"_",wk,"_",treat)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}
#DOWN
#set up data frame 
df_DN=data.frame(row.names="genes")
num=numeric()
#set up loops 
for(ct in celltypes){
  for(cond in conditions){
    for(wk in weeks){
      for(treat in treatment){
        tryCatch({
#read file
x_DN <- read.csv(paste0(ct,"_",cond,"_",wk,"_",treat,"_SIG_DN_DEG.csv"))
#count genes/rows
num <- nrow(x_DN)
#bind to data frame 
df_DN=cbind(df_DN,num)
#set colnames 
colnames(df_DN)[ncol(df_DN)]=paste0(ct,"_",cond,"_",wk,"_",treat)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}
#merge dataframes, add up and down genes 
# add rownames as a column in each data.frame and bind rows
df <- bind_rows(df_UP %>% add_rownames(), 
          df_DN%>% add_rownames()) %>% 
    # evaluate following calls for each value in the rowname column
    group_by(rowname) %>% 
    # add all non-grouping variables
   summarise_all(sum)
write.csv(df, "DEG_count.csv")
```

GENESETS
```{r}
#see SourceTable1.Genesets for more information on sources 

#GO Cytosolic Ribosomal
ribo <-c("APOD","DDX3X","DHX29","EIF2A","EIF2AK2","EIF2AK3","EIF2AK4","EIF2D","FAU","GM3550","GM6096","GM6133","GM6525","GM14214","GM15013","GM20509","GM45713","GM49804","HBA-A1","HBA-A2","ISG15","LARP4","MCTS1","MRPL1","MRPS11","NUFIP1","PPARGC1A","RACK1","REPIN1","RPL3","RPL3L","RPL4","RPL5","RPL6","RPL7","RPL7A","RPL7L1","RPL8","RPL9","RPL10","RPL10A","RPL10L","RPL11","RPL12","RPL13","RPL13A","RPL14","RPL15","RPL17","RPL18","RPL18A","RPL19","RPL21","RPL22","RPL23","RPL23A","RPL24","RPL26","RPL27","RPL27A","RPL28","RPL29","RPL30","RPL31","RPL32","RPL32L","RPL34","RPL35","RPL35A","RPL36","RPL36A","RPL37","RPL37A","RPL38","RPL39","RPL39L","RPLP0","RPLP1","RPLP2","RPS2","RPS3","RPS3A1","RPS4L","RPS4X","RPS5","RPS6","RPS7","RPS8","RPS9","RPS10","RPS11","RPS12","RPS13","RPS14","RPS15","RPS15A","RPS16","RPS17","RPS18","RPS19","RPS20","RPS21","RPS23","RPS24","RPS25","RPS26","RPS27","RPS27A","RPS27L","RPS28","RPS29","RPSA","RSL24D1","UBA52","ZCCHC17")
ribo = ribo [ribo %in% rownames(combined@assays$RNA@data)]

#AD Associated Astrocyte Signature 
DAA_FIG2<- c("GFAP","APOE","CLU","CST3","MT1","CPE","ID3","CD9","CD81","MT2","FXYD1","CKB","VIM","PRDX6","CSMD1","DBI","CTSB","FTH1","HSD17B4","SLC1A3","CST3","NCAM2","KCNIP4","LUZP2","PLCE1","GRM5","ENOX1","VIM","CADM2")
DAA_FIG2 = DAA_FIG2[DAA_FIG2 %in% rownames(combined@assays$RNA@data)]

#A1 Toxic Astrocyte Signature
ZAMANIAN_A1_AST <- c("SULF2","CRISPLD2","DCN","AMIGO2","FBL5","NFASC","SORBS1","SRGN","TLR2","SAA3","ZC3HAV1","TRIM30A","TGTP1","IRGM1","TNFAIP2","TAPBP","TAPBPL","B2M","H2-D1","H2-K1","H2-Q6","H2-T10","H2-T23","PMSB8","PSMB9","KCTD1","SLC22A4","SLC1A5","IL1R1","IL13RA","IFI44","IGTP","GBP2","PARP14","XAF1","C1RB","C1RA","C4B","C1S","FKBP5","HSP1","MAP3K6","OLFM1","SEMA4C","GPX3","GSR","ACSL5","LY6E","UGT1A1","ENDOU","PLIN4","TSPO","ANGPT1","TGM2","SLC43A3","FBLN5","UGT1A","PSMB8","GBP10","GM3579")
ZAMANIAN_A1_AST = ZAMANIAN_A1_AST[ZAMANIAN_A1_AST %in% rownames(combined@assays$RNA@data)]

#A2 Protective Astroyte Signature 
ZAMANIAN_A2_AST <- c("CD24A","JUB","COL6A1","IGFBP3","COL6A2","COL12A1","LGALS1","PVR","THBS1","TGFBI","VCAN","DPYSL3","ECM1","ADAM12","ADAMTS4","ADAMTS5","LGALS3","NAV2","FGL2","NETO2","OLFML3","AKR1B8","BCAT1","ODC1","ESD","PDE3B","ASNS","CYP1B1","B3GNT5","GCNT2","CTPS","UCK2","CDBPD","KLF5","KLF6","TGIF1","AHR","FOSL1","HMGA","FOSL2","SBNO2","S100A6","ZWINT","CDT1","CCND1","CDK6","GADD45A","HMGA2","EMP1","BDKRB2","GADD45B","SPHK1","SOCS3","CAMK2D","ODZ2","SHISA6","NUPR1","THBD","TLR4","PLP2","CLCF1","CCL2","IL6","LIF","ARPC1B","ACTN1","FSCN1","FLNC","FLNA","MSN","STEAP1","SLC5A3","AKAP12","LMNA","NES","TUBA1A","TUBB6","AGPAT9","CAV1","LASS6","PLA2G4A","CH25H","PTGS2","SERPINE1","ANXA1","ANXA2","ANXA7","ANXA3","S100A11","GPX1","HMOX1","SRXN1","TXNRD1","GCH1","MCL1","LITAF","EDA2R","SULF1","IFI203","PAPPA","PRSS23","LONRF1","VGF","CACNG5","SYT4","NHP2","NOP58","EIF1A","PCBP3","IL13RA1","IL6RA","MET","TNFRSF12A","BDNF","CTGF","GDF15","RNF125","RNF19B","OCIAD2","MRPS6","MTHFD2","EII2","FAM129B","GRB10","TMEM74","KLHDC8A","LRRC59","MEST","TM4SF1","LRRFIP1","STX11","SPATA13","SLC44A3","SLC7A1","AHNAK","AHNAK2","H19","BTG3","CHAC1","HMGA1")
ZAMANIAN_A2_AST = ZAMANIAN_A2_AST[ZAMANIAN_A2_AST %in% rownames(combined@assays$RNA@data)]

#HALLMARK TNFA SIGNALING VIA NFKB
tnf<- c("ABCA1","ACKR3","AREG","ATF3","ATP2B1","B4GALT1","B4GALT5","BCL2A1","BCL3","BCL6","BHLHE40","BIRC2","BIRC3","BMP2","BTG1","BTG2","BTG3","CCL2","CCL20","CCL4","CCL5","CCN1","CCND1","CCNL1","CCRL2","CD44","CD69","CD80","CD83","CDKN1A","CEBPB","CEBPD","CFLAR","CLCF1","CSF1","CSF2","CXCL1","CXCL10","CXCL11","CXCL2","CXCL3","CXCL6","DDX58","DENND5A","DNAJB4","DRAM1","DUSP1","DUSP2","DUSP4","DUSP5","EDN1","EFNA1","EGR1","EGR2","EGR3","EHD1","EIF1","ETS2","F2RL1","F3","FJX1","FOS","FOSB","FOSL1","FOSL2","FUT4","G0S2","GADD45A","GADD45B","GCH1","GEM","GFPT2","GPR183","HBEGF","HES1","ICAM1","ICOSLG","ID2","IER2","IER3","IER5","IFIH1","IFIT2","IFNGR2","IL12B","IL15RA","IL18","IL1A","IL1B","IL23A","IL6","IL6ST","IL7R","INHBA","IRF1","IRS2","JAG1","JUN","JUNB","KDM6B","KLF10","KLF2","KLF4","KLF6","KLF9","KYNU","LAMB3","LDLR","LIF","LITAF","MAFF","MAP2K3","MAP3K8","MARCKS","MCL1","MSC","MXD1","MYC","NAMPT","NFAT5","NFE2L2","NFIL3","NFKB1","NFKB2","NFKBIA","NFKBIE","NINJ1","NR4A1","NR4A2","NR4A3","OLR1","PANX1","PDE4B","PDLIM5","PER1","PFKFB3","PHLDA1","PHLDA2","PLAU","PLAUR","PLEK","PLK2","PLPP3","PMEPA1","PNRC1","PPP1R15A","PTGER4","PTGS2","PTPRE","PTX3","RCAN1","REL","RELA","RELB","RHOB","RIPK2","RNF19B","SAT1","SDC4","SERPINB2","SERPINB8","SERPINE1","SGK1","SIK1","SLC16A6","SLC2A3","SLC2A6","SMAD3","SNN","SOCS3","SOD2","SPHK1","SPSB1","SQSTM1","STAT5A","TANK","TAP1","TGIF1","TIPARP","TLR2","TNC","TNF","TNFAIP2","TNFAIP3","TNFAIP6","TNFAIP8","TNFRSF9","TNFSF9","TNIP1","TNIP2","TRAF1","TRIB1","TRIP10","TSC22D1","TUBB2A","VEGFA","YRDC","ZBTB10","ZC3H12A","ZFP36")
tnf= tnf[tnf%in% rownames(combined@assays$RNA@data)]

#Reactome_HSF1 Pathway 
HSF1 <- c("COL4A6","CRYBA4","DEDD2","DNAJB1","DNAJB6","EEF1A1","FKBP4","GML","HDAC6","HSBP1","HSF1","HSP90AA1","HSP90AB1","HSPA1A","HSPA1B","HSPA1L","HSPA6","HSPB1","HSPB2","HSPH1","MRPL18","PTGES3","RLN1","RPA1","RPA2","RPA3","SERPINH1","TNFRSF21","UBB","VCP","YWHAE")
HSF1 = HSF1[HSF1 %in% rownames(combined@assays$RNA@data)]

#AD Neuron Upregulated Signature (Grubman)
GRUBMAN <- c("QDPR","HSP90AA1","NDRG2","MT3","PSAP","HIF3A","GPRC5B","HSP90AB1","CNTN2","CERCAM","BIN1","BCYRN1","CLK1","DNAJB1","NDRG1")
GRUBMAN = GRUBMAN[GRUBMAN %in% rownames(combined@assays$RNA@data)]

#AD GWAS

GWAS2 <-c("APOE","SORT1","CR1","ADAM17","PRKD3","NCK2","BIN1","WDR12","INPP5D","MME","IDUA","CLNK","RHOH","ANKH","COX7C","TNIP1","RASGEF1C","UNC5CL","TREML2","CD2AP","HS3ST5","UMAD1","ICA1","TMEM106B","JAZF1","EPDR1","SEC61G","SPDYE3","EPHA1","CTSB","PTK2B","CLU","SHARPIN","ABCA1","USP6NL","ANK3","TSPAN14","PLEKHA1","SPI1","MS4A4A","EED","SORL1","TPCN1","FERMT2","SLC24A4","SPPL2A","MINDY2","APH1B","SNX1","CTSH","DOC2A","BCKDK","IL34","MAF","PLCG2","FOXF1","WDR81","SCIMP","MYO15A","GRN","WNT3","ABI3","TSPOAP1","ACE","ABCA7","KLF16","SIGLEC11","LILRB2","RBCK1","CASS4","SLC2A4RG","APP","ADAMTS1")

```

GWAS HEATMAP
```{r}
genes <- GWAS2
path <- "AD GWAS GENE EXPRESSION"
#toggle
#cells <- c("ASC")
#cells <- c("IN_NEU")
cells <- c("EX_NEU")
conds <- c("ctr_v_w3","otau_v_w3")
#conds <- c("EX_NEU_ctr_w3_v","EX_NEU_otau_w3_v","IN_NEU_ctr_w3_v","IN_NEU_otau_w3_v","ASC_ctr_w3_v","ASC_otau_w3_v")
#colnames <- c("OUD-","OUD-")
colwidth <- 5
heights <- 10
#fig <- "5"

#START
Nuclei <- combined

Nuclei=SetIdent(Nuclei,value="condition.treat.week")
cond=conds
temp=subset(Nuclei,idents=cond)

temp=SetIdent(temp,value="celltype")
celltype=cells
temp=subset(temp,idents=celltype)

genes = genes[genes %in% rownames(temp@assays$RNA@data)]

temp=SetIdent(temp,value="condition.treat.week")
temp=FindVariableFeatures(temp)
temp=ScaleData(temp,vars.to.regress="nCount_RNA",features=unique(c(temp@assays$RNA@var.features,genes)))

heatcolor=brewer.pal(11,"RdYlBu")

what=AverageExpression(temp,features=genes,verbose=F,slot="scale.data")$RNA
min=quantile(unlist(what),0.1)
mid=quantile(unlist(what),0.5)
max=quantile(unlist(what),0.9)

p <- Heatmap(what,
             width = unit(colwidth, "cm"),
             name="Gene Expression",
             col=colorRamp2(c(min, 0, max), c("blue", "white", "red")),
             clustering_method_columns = "ward.D2",
             cluster_rows=FALSE,
             cluster_columns=FALSE,
             column_split=do.call("rbind",strsplit(colnames(what),split="_"))[,1],
             show_column_names=T,
             show_row_names=F,
             rect_gp = gpar(col = "black", lwd = 2), 
             show_column_dend = FALSE, 
             show_row_dend = FALSE, 
             column_labels = c("-","+"), 
             column_names_rot = 0, 
             column_names_centered = T,
             column_title = celltype, 
             column_title_gp = gpar(fontface="bold"), 
             column_names_gp = gpar(fontface="bold"), 
             heatmap_legend_param = list(direction = "horizontal"),
             row_names_gp = gpar(fontface="bold"))


p<- ComplexHeatmap::draw(p, heatmap_legend_side="bottom")

pdf(paste0("./gwas/",celltype,"_gwas_heatmap.pdf"),width = 8, height = heights)
p
dev.off()


```

SUBSET ANALYSIS
```{r}
#subset to main conditions
Idents(combined)<-"condition.treat.week"
subset <- subset(combined, idents=c("ctr_v_w3", "otau_v_w3"))
Idents(subset)<-"condition.treat.week"

CELLTYPES <- c("ASC")

#CELLTYPES <- c("IN_NEU","EX_NEU")

for (CTS in CELLTYPES){
   tryCatch({

Idents(subset)<-"celltype"
subset_ct <- subset(subset, idents=CTS)
subset_ct <- FindVariableFeatures(subset_ct, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(subset_ct)
subset_ct <- ScaleData(subset_ct, features = all.genes)
subset_ct <- RunPCA(subset_ct, verbose = FALSE)
subset_ct <- RunUMAP(subset_ct, reduction = "pca", dims = 1:15)
subset_ct <- FindNeighbors(subset_ct, dims = 1:15, verbose = FALSE)
subset_ct <- FindClusters(subset_ct, verbose = FALSE)

#Save seurat object 
save(subset_ct,file="subset_ASC.RData")

 }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}


#relevel
#Condition
Idents(subset_ct) <- "condition.treat.week"
# Define an order of cluster identities
condition_levels <-  c("ctr_v_w3", "otau_v_w3")
# Relevel object@ident
subset_ct@meta.data$condition.treat.week <- factor(x = subset_ct@meta.data$condition.treat.week, levels = condition_levels)

#colors
cond_color_sub <- c(col_ctr_v_w3,col_otau_v_w3)
cond_label  <- c("Control", "AstTau")

#umaps
pdf(paste0("./subset/",CTS,"/",CTS,"_cluster_UMAP.pdf"))
DimPlot(subset_ct, reduction = "umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE, label.size = 8, raster = F)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15) )
dev.off()

pdf(paste0("./subset/",CTS,"/",CTS,"_condition_UMAP.pdf"))
DimPlot(subset_ct, reduction = "umap", group.by = "condition", shuffle = TRUE, raster=F)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15))+scale_color_manual(labels= cond_label,values= cond_color_sub)
dev.off()

#combined clusters 

#celltype assignment
Idents(subset_ct)<- subset_ct$seurat_clusters
subset_ct<- RenameIdents(subset_ct, `0` = "A", `1` = "B", `2` = "C", `3` = "B", `4` = "D", `5` = "C", `6` = "E", `7` = "F", `8` ="G", `9` = "C", `10` = "E", `11` ="G", `12` = "F")
subset_ct$cluster2 <- Idents(subset_ct)

#new metadata columns
subset_ct$cluster_condition <- paste(subset_ct$seurat_clusters, subset_ct$condition, sep = "_")
subset_ct$cluster2_condition <- paste(subset_ct$cluster2, subset_ct$condition, sep = "_")

pdf(paste0("./subset/",CTS,"/",CTS,"_cluster_UMAP_2.pdf"))
DimPlot(subset_ct, reduction = "umap", group.by = "cluster2", label = TRUE, repel = TRUE, label.size = 8, raster = F)+theme(text = element_text(face = "bold"),plot.title = element_text(size = 0), axis.text = element_text(size = 15) )
dev.off()


#cluster count barplot

#colored by condition
Idents(subset_ct) <- "condition"
#table data 
t <- table(subset_ct@meta.data$condition, subset_ct@meta.data$cluster2)
rownames(t) <- c("Control","AstTau")
#convert to percentage of total
t <- prop.table(t,margin = 1)
t <- t*100
#document
write_csv(as.data.frame(t), paste0("./subset/",CTS,"/",CTS,"_cluster_counts_2.csv"))
# melt the data frame for plotting
data.m <- melt(t, id.vars='cluster2')
#plot
plot<- ggplot(data.m, aes(y = Var1, x = value, fill = as.character(Var2))) +geom_bar(stat = 'identity', width = 1,colour="black") +theme(legend.text = element_text(size = 10, colour = "black"), axis.text.y = element_text(size = 10, colour = "black", angle = 20), axis.title.y = element_text(size = 0), axis.title.x = element_text(size = 0),axis.text.x = element_text(face="bold", size = 10),legend.position = "bottom", legend.direction = "horizontal")+ theme(panel.background = element_blank())+ 
  labs(fill='Cluster')+theme(aspect.ratio = .2)+ guides(fill = guide_legend(nrow = 1))

pdf(paste0("./subset/",CTS,"/",CTS,"_cluster_barplot_2.pdf"))
print(plot)
dev.off()

#cluster markers
Idents(subset_ct)<-"cluster2"
markers <- FindAllMarkers(subset_ct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)

write_csv(markers, paste0("./subset/",CTS,"/",CTS,"_cluster_markers_sub_2.csv"))

markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10

pdf(paste0("./subset/",CTS,"/",CTS,"_cluster_heatmap_2.pdf"))
print(DoHeatmap(subset_ct, features = top10$gene ) + NoLegend()+theme(axis.text.y = element_text(size=5)))+ scale_fill_gradientn(colors = c("blue", "white", "red"))
dev.off()

```

DEG HEATMAP
```{r}
#make top 25 gene list 

##TOGGLE to output +/- PU 
##V
treatment <- c("v")
weeks <- c("w1","w2","w3")
colnames <- c("","","")
label <- "v"
rowwidth<- 25
docheight <-11
##PU
# treatment <- c("v","pu")
# weeks <- c("w1","w3")
# colnames <- c("","","")
# label<-"pu"
# rowwidth<- 25
# docheight <-11

##TOGGLE to output individual plot for each cell type 
celltypes <- c("EX_NEU")
#celltypes <- c("IN_NEU")
#celltypes <- c("ASC")

conditions <- c("otau")
direction <- c("UP","DN")

gs=character()

for(cond in conditions){
for(dir in direction){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
        tryCatch({
#selecting top 25 genes 
data <- read.csv2(paste0(ct,"_",cond,"_",wk,"_",trt,"_","SIG_",dir,"_DEG.csv"))
data2 <- head(data, 25)
data3<- data2$name
gs=c(gs,data3)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
       }}}}}
genes=unique(gs)

#sest up complexheatmap dataframe 
gs=genes
df=data.frame(row.names=gs)
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
        tryCatch({
res=rep(0,length(gs))
ressig=rep(0,length(gs))
data <- read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_DEG.csv"))
rownames(data)=data$name
res[gs %in% data$name]=data[gs[gs %in% data$name],"avg_log2FC"]
df=cbind(df,res)
colnames(df)[ncol(df)]=paste0(ct,"_",cond,"_",wk,"_",trt)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
   }}}}

#set heatmap formatting 
ht_opt(legend_border = "black",heatmap_border = TRUE)
colwidth<- 2

#color code row names based on ontology pathway 
#1= HSP 
celltype1 <-HSF1 
selectcolor1 <- "red"
#2= TNF
celltype2 <-tnf
selectcolor2 <- "purple"
#3= RIBO 
celltype3 <-ribo
selectcolor3 <- "green"
keyvals <- ifelse(rownames(df) %in% celltype1, selectcolor1,
      ifelse(rownames(df) %in% celltype2, selectcolor2,
             ifelse(rownames(df) %in% celltype3, selectcolor3,
       "black")))
#make heatmap
ht<- Heatmap(df,width = unit(colwidth, "cm"), height = unit(rowwidth, "cm"), row_names_gp = gpar(fontsize = 6, face="bold",col = keyvals),column_names_gp = gpar(fontsize=8, face="bold"), cluster_columns = F, cluster_rows = F,column_labels = colnames, column_names_rot=0, column_names_centered = TRUE, column_names_side = "top",name = "log2FC")
pdf(paste0(ct,"_",label,"_", dir,"_DEG_heatmap.pdf"), width = 2, height = docheight)
ht
dev.off()
```

GPRO HEATMAP CSV
```{r}
#create custom GMT ID code
# upload_GMT_file(".gmt")
# "gp__zSEF_sD9Q_d1M"- all mSIGdB HALLMARK,C2,and C5 pathways 
#create custom domain_scope, the full list of genes in your assay as the background 
total.genes <- rownames(combined@assays$RNA@data)
write.csv(total.genes, "total_genes.csv")
#loading in data for loop, see the 'for' and read.csv arguments
celltypes <- c("EX_NEU","IN_NEU","ASC")
conditions <- c("otau","asyn")
weeks <- c("w1","w2","w3")
treatment <- c("v","pu")
direction=c("SIG_DN_DEG","SIG_UP_DEG")
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
      for(dir in direction){
     tryCatch({
x <- read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_",dir,".csv"))
x <- as.character(x$name)
#GPRO
gostres <- gost(query = x, 
                organism = "gp__zSEF_sD9Q_d1M", ordered_query = TRUE, 
                multi_query = FALSE, significant = FALSE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = FALSE, 
                user_threshold = 0.05, correction_method = "fdr", 
                domain_scope = "custom", custom_bg = total.genes, 
                numeric_ns = "", sources = "GO", as_short_link = FALSE)
gostres2 <- gostres$result 
gostres3 <- gostres2[1:13]
#write.csv 
write_csv(gostres3, paste0(ct,"_",cond,"_",wk,"_",trt,"_",dir,"_gpro.csv"))
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}}
```

GPRO GEM TXT 
```{r}
#create custom GMT ID code 
#upload_GMT_file(".gmt")
# "gp__zSEF_sD9Q_d1M"- all mSIGdB HALLMARK,C2,and C5 pathways
#create custom domain_scope, the full list of genes in your assay as the background 
total.genes <- rownames(combined@assays$RNA@data)
write.csv(total.genes, "total_genes.csv")
#loading in data for loop, see the 'for' and read.csv arguments
celltypes <- c("EX_NEU","IN_NEU","ASC")
conditions <- c("otau","asyn")
weeks <- c("w1","w2","w3")
treatment <- c("v","pu")
direction=c("SIG_DN_DEG","SIG_UP_DEG")
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
      for(dir in direction){
     tryCatch({
x <- read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_",dir,".csv"))
x <- as.character(x$name)
#GPRO
gostres <- gost(query = x, 
                organism = "gp__zSEF_sD9Q_d1M", ordered_query = TRUE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = TRUE, 
                user_threshold = 0.05, correction_method = "fdr", 
                domain_scope = "custom", custom_bg = total.genes, 
                numeric_ns = "", sources = "GO", as_short_link = FALSE)
gostres$result=gostres$result[gostres$result$intersection_size>1,]
gem <- gostres$result[,c("term_id", "term_name", "p_value", "intersection")]
colnames(gem) <- c("GO.ID", "Description", "p.Val", "Genes")
gem$FDR <- gem$p.Val
gem$Phenotype = "+1"
gem <- gem[,c("GO.ID", "Description", "p.Val", "FDR", "Phenotype", "Genes")]
toMatch <- c("^GO_", "^REACTOME_", "^KEGG_","^HALLMARK_")
gem=gem[grep(paste0(toMatch,collapse="|"),gem$GO.ID),]
write.table(gem, file = paste0(ct,"_",cond,"_",wk,"_",trt,"_",dir,"_gpro_gem.txt"), sep = "\t", quote = F, row.names = F)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}}
```

GPRO GS W/OTAU ONLY
```{r}
celltypes <- c("EX_NEU","IN_NEU","ASC")
conditions <- c("otau")
weeks <- c("w1","w2","w3")
treatment <- c("v")
gs=character()
toMatch <- c("^GO_", "^REACTOME_", "^KEGG_","^HALLMARK_")
#UP
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
   tryCatch({
t=read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_","SIG_UP_DEG_gpro.csv"))
t=t[grep(paste0(toMatch,collapse="|"),t$term_id),]
t$term_id=as.character(t$term_id)
t=t[t$p_value<0.05,]
t=t[t$intersection_size>1,]
t=t[t$term_size<500,]
t <-t[order(t$p_value),]
gs=c(gs,head(t$term_id,25))
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}
gs=unique(gs)
#save rds
saveRDS(gs, file = paste0("UP_GPRO_GS.rds"))
write_csv(as.data.frame(gs),paste0("UP_GPRO_GS.csv"))
#DN
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
    tryCatch({
t=read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_","SIG_DN_DEG_gpro.csv"))
t=t[grep(paste0(toMatch,collapse="|"),t$term_id),]
t$term_id=as.character(t$term_id)
t=t[t$p_value<0.05,]
t=t[t$intersection_size>1,]
t=t[t$term_size<500,]
t <-t[order(t$p_value),]
gs=c(gs,head(t$term_id,25))
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}
gs=unique(gs)
#save rds
saveRDS(gs, file = paste0("DN_GPRO_GS.rds"))
write_csv(as.data.frame(gs),paste0("DN_GPRO_GS.csv"))
```

GPRO HEATMAP W/ OTAU V ONLY
```{r}
#colors
heatcolor1=brewer.pal(11,"RdYlBu")[1]
heatcolor2=brewer.pal(11,"RdYlBu")[5]
cellcol<- c(col_EX_NEU, col_IN_NEU,col_ASC)
#samples
celltypes <- c("EX_NEU","IN_NEU","ASC")
conditions <- c("otau")
weeks <- c("w1","w2","w3")
treatment <- c("v")
#UP
gs=readRDS(paste0("UP_GPRO_GS.rds"))
df=data.frame(row.names=gs)
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
     tryCatch({
res=rep(0,length(gs))
ressig=rep(0,length(gs))
t=read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_","SIG_UP_DEG_gpro.csv"))
t <- na.omit(t)
t$term_id=as.character(t$term_id)
rownames(t)=t$term_id
res[gs %in% t$term_id]=t[gs[gs %in% t$term_id],"p_value"]
df=cbind(df,res)
colnames(df)[ncol(df)]=paste0(ct,"_",cond,"_",wk,"_",trt)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

}}}}
log.df <- -log10(df)
log.df[log.df == Inf] <- 0
#shorten rownames
row.names(log.df)<-gsub("_", " ", row.names(log.df), fixed=TRUE)
row.names(log.df) <- strtrim(row.names(log.df), width=50)
#set split
split = rep(1:length(celltypes), each = 3)
#set colnames
colnames <- rep(c("7","14","21"), times = length(celltypes))
#set annotations
ha = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = cellcol), labels = c( celltypes <- c("EX_NEU","IN_NEU","ASC")), labels_gp = gpar(fontsize=15)))
#make heatmap
ht<- Heatmap(log.df,
        name = "-log10(p-value)",
        column_names_rot=0,
        column_names_centered = TRUE,
        row_names_gp = gpar(fontsize = 4),
        column_labels = colnames,
        column_names_side = "top",
        column_names_gp = gpar(fontsize=12),
        cluster_columns=F,
        cluster_rows=F,
        show_row_dend = F,
        col=colorRamp2(c(0,1.30102,1.30103,max(log.df)),c("white","white",heatcolor2, heatcolor1)),
        column_split = split,
        top_annotation = ha, 
        column_title = NULL,
        heatmap_legend_param = list(direction = "horizontal"),
       row_names_max_width = max_text_width( rownames(log.df), gp = gpar(fontsize = 4, face='bold')))
ht<- ComplexHeatmap::draw(ht, heatmap_legend_side="bottom")
pdf(paste0("SIG_UP_DEG_GPRO_heatmap.pdf"))
ht
dev.off()
```

GPRO HEATMAP W/ OTAU V+PU 
```{r}
#colors
heatcolor1=brewer.pal(11,"RdYlBu")[1]
heatcolor2=brewer.pal(11,"RdYlBu")[5]
cellcol<- c(col_EX_NEU, col_IN_NEU,col_ASC)
#samples
celltypes <- c("EX_NEU","IN_NEU","ASC")
conditions <- c("otau")
weeks <- c("w1","w2","w3")
treatment <- c("v","pu")
#UP
gs=readRDS(paste0("UP_GPRO_GS.rds"))
df=data.frame(row.names=gs)
for(cond in conditions){
for(ct in celltypes){
  for(wk in weeks){
   for(trt in treatment){
   tryCatch({
res=rep(0,length(gs))
ressig=rep(0,length(gs))
t=read.csv(paste0(ct,"_",cond,"_",wk,"_",trt,"_","SIG_UP_DEG_gpro.csv"))
t <- na.omit(t)
t$term_id=as.character(t$term_id)
rownames(t)=t$term_id
res[gs %in% t$term_id]=t[gs[gs %in% t$term_id],"p_value"]
df=cbind(df,res)
colnames(df)[ncol(df)]=paste0(ct,"_",cond,"_",wk,"_",trt)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}}}
log.df <- -log10(df)
log.df[log.df == Inf] <- 0
#shorten rownames
row.names(log.df)<-gsub("_", " ", row.names(log.df), fixed=TRUE)
row.names(log.df) <- strtrim(row.names(log.df), width=50)
#set split
split = rep(1:length(celltypes), each = 5)
#set colnames
colnames <- rep(c("7","14","14","21","21"), times = length(celltypes))
#set annotations
ha = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = cellcol), labels = celltypes, labels_gp = gpar(fontsize=15)))
#heatmap
ht<- Heatmap(log.df,
        name = "-log10(p-value)",
        column_names_rot=0,
        column_names_centered = TRUE,
        row_names_gp = gpar(fontsize = 4),
        column_labels = colnames,
        column_names_side = "top",
        column_names_gp = gpar(fontsize=12),
        cluster_columns=F,
        cluster_rows=F,
        show_row_dend = F,
        col=colorRamp2(c(0,1.30102,1.30103,max(log.df)),c("white","white",heatcolor2, heatcolor1)),
        column_split = split,
        top_annotation = ha, 
        column_title = NULL,
        heatmap_legend_param = list(direction = "horizontal"),
       row_names_max_width = max_text_width( rownames(log.df), gp = gpar(fontsize = 4, face='bold')))
ht<- ComplexHeatmap::draw(ht, heatmap_legend_side="bottom")
pdf(paste0("PU_SIG_UP_DEG_GPRO_heatmap.pdf"))
ht
dev.off()
```

LITERATURE HEATMAP COMPARISON
```{r}
#gather genesets from literature to compare 
##AstTAU, common upregulated genes across IN_NEU, EX_NEU, ASC, 11 genes, p<0.05, log2FC < 0.25
csv_1 <- read.csv(paste0("IN_NEU_otau_w3_v_SIG_UP_DEG.csv"))
csv_2 <- read.csv(paste0("EX_NEU_otau_w3_v_SIG_UP_DEG.csv"))
csv_3 <- read.csv(paste0("ASC_otau_w3_v_SIG_UP_DEG.csv"))
csv <- merge(csv_1,csv_2, by = 1, all=TRUE)
csv2 <- merge(csv,csv_3, by = 1, all=FALSE)
write_csv(csv2, paste0("asteroid_gprocomp.csv"))
##AstaSYN, common upregulated genes across IN_NEU, EX_NEU, ASC, 491 genes, p<0.05, log2FC < 0.25
csv_1 <- read.csv(paste0("IN_NEU_asyn_w3_v_SIG_UP_DEG.csv"))
csv_2 <- read.csv(paste0("EX_NEU_asyn_w3_v_SIG_UP_DEG.csv"))
csv_3 <- read.csv(paste0("ASC_asyn_w3_v_SIG_UP_DEG.csv"))
csv <- merge(csv_1,csv_2, by = 1, all=TRUE)
csv2 <- merge(csv,csv_3, by = 1, all=FALSE)
write_csv(csv2, paste0("x_asteroid_asyn_gprocomp.csv"))
##MATHYS common upregulated genes across AST, IN, EX  in Figure 2c Late versus early pathology,  FDR < 0.01, log2FC > 0.25, 16 genes 
mathys <- c("HSPA1A","GAPDH","DYNLL1","UBB","FTH1","CRYAB","PTMA","UBC","HSPB1","NDUFA4","STMN1","PCSK1N","COX7C","SH3BFRL3","HSP90AA1","HSPA8")
mathys <- as.data.frame(mathys)
write_csv(mathys, paste0("mathys_gprocomp.csv"))
##GRUBMAN genes upregulated in Neuron and Astrocyte (DEG cluster 5 and 7, source data), absolute log fold change > 1 and FDR < 0.01 
grubman <- c("BOK","CRYAB","DNAJB2","FKBP5","GFAP","HSPA1A","HSPB1","LINC00486","LINGO1","MT-ATP6","MT-CO2","MT-CO3","MT-CYB","MT-ND1","MT-ND2","MT-ND3","MT-ND4","NEAT1","SPP1","ABCC8","ABLIM1","ADAMTS18","ADCK3","AEBP1","AHNAK","ANLN","ATF4","B3GAT3","BCL6","BCOR","BCYRN1","BIN1","BRSK2","CD163L1","CDK10","CERCAM","CH507-513H4.1","CHD5","CHI3L1","CLK1","CLSTN1","CNTN2","CPSF3L","CRELD1","CRISPLD2","CSPG4","CSRP1","DDB1","DHCR24","DNAJA1","DNAJB1","DPP7","EEF2","EEF2K","FKBP4","GIT1","GPRC5B","GRIN1","HIF3A","HP1BP3","HSP90AA1","HSP90AB1","HSPA1B","HSPH1","ITGB4","KIF19","LAMTOR4","LGI3","LIFR","LSS","MAPK8IP1","MGLL","MICALL2","MT3","MYLPF","NDRG1","NDRG2","NUMA1","PABPN1","PAPOLA","PHYHIP","PLEC","PLOD3","PRNP","PSAP","PTGES3","QDPR","RANGAP1","RAPGEF3","RHPN1","RP3-406A7.7","SEMA3E","SLC26A3","SLC38A2","SLC5A11","SNPH","SRSF6","STIP1","TEF","TMEM259","TMEM63B","TPPP","TSPYL2","TTYH2","TYRO3","WASH1","ZC3H11A","ZFAND5","ZNF692")
grubman <- as.data.frame(grubman)
write_csv(grubman, paste0("grubman_gprocomp.csv"))
##VELMESHEV negative control data set, Fig. 2 a,b highlighted upregulated genes in neuronal and non-neuronal cell types, 22 genes 
Velmeshev <- c("CHSY3","GLRA3","NEGR1","FSTL5","CDH2","SSBP2","GPATCH8","GABRB1","CH17-472G23.1","RP11-444D3.1","NSD1","GPHN","PACS1","GSK3B","PLEKHG1","PCSK5","RUNDC3B","FBLN1","FARP1","BPTF","USP53","FAF1")
Velmeshev <- as.data.frame(Velmeshev)
write_csv(Velmeshev, paste0("velmeshev_gprocomp.csv"))
#run gpro from CSV
csv_files <- list.files (pattern    = "*gprocomp.csv", full.names = T)
for(file in csv_files){
x<- read.csv(paste0(file), header = FALSE)
x <- as.character(x$V1)
gostres <- gost(query = x, 
                organism = "gp__zSEF_sD9Q_d1M", ordered_query = FALSE, 
                multi_query = FALSE, significant = FALSE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = FALSE, 
                user_threshold = 0.05, correction_method = "fdr", 
                numeric_ns = "", sources = "GO", as_short_link = FALSE)
gostres2 <- gostres$result 
gostres3 <- gostres2[1:13]
#write.csv 
write_csv(gostres3, paste0(file,"_gpro.csv"))
}
#gs list 
csv_files <- list.files (pattern   = "*gprocomp.csv_gpro.csv", full.names = T)
csv_files <- csv_files[1:3]
toMatch <- c("^GO_", "^REACTOME_", "^KEGG_","^HALLMARK_")
gs=character()
for(file in csv_files){
t<- read.csv(paste0(file), header = TRUE)
t=t[grep(paste0(toMatch,collapse="|"),t$term_id),]
t$term_id=as.character(t$term_id)
t=t[t$p_value<0.05,]
t=t[t$intersection_size>2,]
t <-t[order(t$p_value),]
gs=c(gs,head(t$term_id,10))
}
gs=unique(gs)
#save rds
saveRDS(gs, file = paste0("GPROCOMP_GS.rds"))
write_csv(as.data.frame(gs),paste0("GPROCOMP_GS.csv"))
#heatmap 
df=data.frame(row.names=gs)
files <- list.files (pattern   = "*gprocomp.csv_gpro.csv", full.names = T)
for(file in files){
res=rep(0,length(gs))
ressig=rep(0,length(gs))
t=read.csv(paste0(file))
t$term_id=as.character(t$term_id)
rownames(t)=t$term_id
res[gs %in% t$term_id]=t[gs[gs %in% t$term_id],"p_value"]
df=cbind(df,res)
colnames(df)[ncol(df)]=paste0(file)
}
log.df <- -log10(df)
log.df[log.df == Inf] <- 0
names(log.df)[1] <- "AstTAU"
names(log.df)[2] <- "Grubman"
names(log.df)[3] <- "Mathys"
names(log.df)[4] <- "Velmeshev"
names(log.df)[5] <- "AstASYN"
#wrapping strings
#replace underscore with space in names
a<-gsub("_", " ", row.names(log.df), fixed=TRUE)
#replace rownames in log.df
row.names(log.df) <- a
#set trim 
row.names(log.df) <- strtrim(row.names(log.df), width=50)
#colors
heatcolor1=brewer.pal(11,"RdYlBu")[2]
heatcolor2=brewer.pal(11,"RdYlBu")[5]
#heatmap
ht<- Heatmap(log.df,
        name = "-log10(p-value)",
        column_names_rot=25,
        column_names_centered = TRUE,
        row_names_gp = gpar(fontsize =8, face= 'bold'),
        column_names_side = "top",
        column_names_gp = gpar(fontsize=12, face='bold'),
        cluster_columns=F,
        cluster_rows=F,
        show_row_dend = FALSE,
        col=colorRamp2(c(0,1.30102,1.30103,max(log.df)),c("white","white",heatcolor2, heatcolor1)),
        row_names_max_width = max_text_width( rownames(log.df), gp = gpar(fontsize = 12, face='bold')),
        heatmap_legend_param = list(direction = "horizontal"))
ht<- ComplexHeatmap::draw(ht, heatmap_legend_side="bottom")
pdf(paste0("comp_gpro_heatmap.pdf"))
ht
dev.off()
```

LIGER COMPARATIVE INTEGRATION 
```{r}
#load datasets for integration 
#AstTau 
load.Rdata("combined_named.RData","x.integrated")
#Badhuri et al., (2020)
so_o <- readRDS("./seurat_stress_organoid.rds") 
#pre-process 
x.integrated.data <- as.SingleCellExperiment(x.integrated)
celseq_o.data <- as.SingleCellExperiment(so_o)
keep_genes <- Reduce(intersect, list(rownames(x.integrated.data),rownames(celseq_o.data)))
x.integrated.data <- x.integrated.data[match(keep_genes, rownames(x.integrated.data)), ]
celseq_o.data <- celseq_o.data[match(keep_genes, rownames(celseq_o.data)), ]
x.integrated.data <- logNormCounts(x.integrated.data)
celseq_o.data <- logNormCounts(celseq_o.data)
#normalize
## x.integrated.data
# Fitting a trend:
library(DelayedMatrixStats)
x.integrated.data.means <- rowMeans(logcounts(x.integrated.data))
x.integrated.data.vars <- rowVars(logcounts(x.integrated.data))
fit_x.integrated.data <- fitTrendVar(x.integrated.data.means, x.integrated.data.vars) 
dec_x.integrated.data <- modelGeneVar(x.integrated.data)
dec_x.integrated.data$Symbol_TENx <- rowData(x.integrated.data)$Symbol_TENx
dec_x.integrated.data <- dec_x.integrated.data[order(dec_x.integrated.data$bio, decreasing = TRUE), ]
## celseq_o.data
celseq_o.data.means <- rowMeans(logcounts(celseq_o.data))
celseq_o.data.vars <- rowVars(logcounts(celseq_o.data))
fit_celseq_o.data <- fitTrendVar(celseq_o.data.means, celseq_o.data.vars) 
dec_celseq_o.data <- modelGeneVar(celseq_o.data)
dec_celseq_o.data$Symbol_TENx <- rowData(celseq_o.data)$Symbol_TENx
dec_celseq_o.data <- dec_celseq_o.data[order(dec_celseq_o.data$bio, decreasing = TRUE), ]
# select the most informative genes that are shared across all datasets:
universe <- Reduce(intersect, list(rownames(dec_x.integrated.data),rownames(dec_celseq_o.data)))
mean.bio <- (dec_x.integrated.data[universe,"bio"] + dec_celseq_o.data[universe,"bio"])/2
hvg_genes <- universe[mean.bio > 0]
#x.integrated colData 
x.integrated.data.ident<-colData(x.integrated.data)$ident
x.integrated.data.sizeFactor<-colData(x.integrated.data)$sizeFactor
colData(x.integrated.data)$ident<-NULL
colData(x.integrated.data)$sizeFactor<-NULL
colData(x.integrated.data)$Cluster<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$Sample<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$Line<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$Protocol<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$Age<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$iPSCorhESC<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$Class<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$State<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$Type<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$Subtype<-colData(x.integrated.data)$orig.ident
colData(x.integrated.data)$ident<-x.integrated.data.ident
colData(x.integrated.data)$sizeFactor<-x.integrated.data.sizeFactor
colData(x.integrated.data)$new.orig.ident<-colData(x.integrated.data)$sample_id
#celseq colData 
celseq_o.data.orig.ident<-colData(celseq_o.data)$orig.ident
colData(celseq_o.data)$orig.ident<-"stress_organoid"
celseq_o.data.Cluster<-colData(celseq_o.data)$Cluster
celseq_o.data.Sample<-colData(celseq_o.data)$Sample
celseq_o.data.Line<-colData(celseq_o.data)$Line
celseq_o.data.Protocol<-colData(celseq_o.data)$Protocol
celseq_o.data.Age<-colData(celseq_o.data)$Age
celseq_o.data.iPSCorhESC<-colData(celseq_o.data)$iPSCorhESC
celseq_o.data.Class<-colData(celseq_o.data)$Class
celseq_o.data.State<-colData(celseq_o.data)$State
celseq_o.data.Type<-colData(celseq_o.data)$Type
celseq_o.data.Subtype<-colData(celseq_o.data)$Subtype
celseq_o.data.ident<-colData(celseq_o.data)$ident
celseq_o.data.sizeFactor<-colData(celseq_o.data)$sizeFactor
colData(celseq_o.data)$Cluster<-NULL
colData(celseq_o.data)$Sample<-NULL
colData(celseq_o.data)$Line<-NULL
colData(celseq_o.data)$Protocol<-NULL
colData(celseq_o.data)$Age<-NULL
colData(celseq_o.data)$iPSCorhESC<-NULL
colData(celseq_o.data)$Class<-NULL
colData(celseq_o.data)$State<-NULL
colData(celseq_o.data)$Type<-NULL
colData(celseq_o.data)$Subtype<-NULL
colData(celseq_o.data)$ident<-NULL
colData(celseq_o.data)$sizeFactor<-NULL
colData(celseq_o.data)$percent_mt<-"stress_organoid"
colData(celseq_o.data)$sample_id<-"stress_organoid"
colData(celseq_o.data)$batch<-"stress_organoid"
colData(celseq_o.data)$condition<-"stress_organoid"
colData(celseq_o.data)$week<-"stress_organoid"
colData(celseq_o.data)$treatment<-"stress_organoid"
colData(celseq_o.data)$batch_id<-"stress_organoid"
colData(celseq_o.data)$integrated_snn_res.0.5<-"stress_organoid"
colData(celseq_o.data)$seurat_clusters<-"stress_organoid"
colData(celseq_o.data)$celltype<-"stress_organoid"
colData(celseq_o.data)$celltype.sample<-"stress_organoid"
colData(celseq_o.data)$celltype.condition<-"stress_organoid"
colData(celseq_o.data)$celltype.batch<-"stress_organoid"
colData(celseq_o.data)$celltype.week<-"stress_organoid"
colData(celseq_o.data)$celltype.treat<-"stress_organoid"
colData(celseq_o.data)$celltype.condition.week.treat<-"stress_organoid"
colData(celseq_o.data)$condition.treat<-"stress_organoid"
colData(celseq_o.data)$condition.treat.week<-"stress_organoid"
colData(celseq_o.data)$Cluster<-celseq_o.data.Cluster
colData(celseq_o.data)$Sample<-celseq_o.data.Sample
colData(celseq_o.data)$Line<-celseq_o.data.Line
colData(celseq_o.data)$Protocol<-celseq_o.data.Protocol
colData(celseq_o.data)$Age<-celseq_o.data.Age
colData(celseq_o.data)$iPSCorhESC<-celseq_o.data.iPSCorhESC
colData(celseq_o.data)$Class<-celseq_o.data.Class
colData(celseq_o.data)$State<-celseq_o.data.State
colData(celseq_o.data)$Type<-celseq_o.data.Type
colData(celseq_o.data)$Subtype<-celseq_o.data.Subtype
colData(celseq_o.data)$ident<-celseq_o.data.ident
colData(celseq_o.data)$sizeFactor<-celseq_o.data.sizeFactor
colData(celseq_o.data)$new.orig.ident<-celseq_o.data.orig.ident
#set colnames 
colnames(colData(x.integrated.data))
colnames(colData(celseq_o.data))
#rm processing files 
rm(dec_x.integrated.data)
rm(dec_celseq_o.data)
rm(fit_x.integrated.data)
rm(fit_celseq_o.data)
rm(x.integrated.data.means)
rm(x.integrated.data.vars)
rm(celseq_o.data.means)
rm(celseq_o.data.vars)
rm(celseq_o.data)
rm(x.integrated.data)
#get counts 
# total raw counts
counts <- cbind(counts(x.integrated.data), counts(celseq_o.data))
# total normalized counts (with multibatch normalization)
logcounts <- cbind(logcounts(x.integrated.data), logcounts(celseq_o.data))
# sce object of the combined data 
sce <- SingleCellExperiment( 
    assays = list(counts = counts, logcounts = logcounts),  
    rowData = rowData(x.integrated.data), # same as rowData(pbmc4k) 
    colData = rbind(colData(x.integrated.data), colData(celseq_o.data)) 
)
# store the hvg_genes from the prior section into the sce objects metadata slot via:
metadata(sce)$hvg_genes <- hvg_genes
sce <- runPCA(sce,ncomponents = 30)
sce.seurat <- as.Seurat(sce)
#rm(sce)
#clustering workflow 
sce.seurat <- NormalizeData(sce.seurat)
sce.seurat <- FindVariableFeatures(sce.seurat)
sce.seurat <- ScaleData(sce.seurat, split.by = "orig.ident", do.center = FALSE)
sce.seurat <- RunOptimizeALS(sce.seurat, k = 20, lambda = 5, split.by = "orig.ident")
sce.seurat <- RunQuantileNorm(sce.seurat, split.by = "orig.ident")
sce.seurat <- FindNeighbors(sce.seurat, reduction = "iNMF", dims = 1:20)
sce.seurat <- FindClusters(sce.seurat, resolution = 0.3)
sce.seurat <- RunUMAP(sce.seurat, dims = 1:ncol(sce.seurat[["iNMF"]]), reduction = "iNMF")
#save intergrated object 
saveRDS(sce.seurat, file = "./x.integrated/x_stress_integrated_seurat.rds")
```

LIGER COMPARATIVE VISUALIZATION 
```{r}
data <- readRDS(paste0("path"))

#downsample
data <- SetIdent(data, value = "orig.ident")
data_sub <- subset(data,downsample=133228)

#visualize split by cluster  
pdf("cerebralorganoid_cluster_umap.pdf")
DimPlot(data_sub, reduction = "umap", label = T, pt.size = 0.2, group.by = "seurat_clusters",  shuffle = T, raster = T, repel = T, label.size = 10)+theme(title = element_text(size = 0)) +theme(text = element_text(face = "bold"))
dev.off()

#visual cell type markers feature plo t
features <- c("TOP2A","NES","GAP43","STMN2","GAD1","SLC17A6","VIM","SLC1A3","GFAP")
pdf("cerebralorganoid_cluster_feature.pdf")
FeaturePlot(data_sub, features = features, order = T)
dev.off()

#celltype assignment
Idents(data_sub)<- data_sub$seurat_clusters
data_sub <- RenameIdents(data_sub, `0` = "NEU_P", `1` = "NEU", `2` = "NEU_P", `3` = "ASC_P", `4` = "ASC_P", `5` = "ASC", `6` = "NEU", `7` = "NEU_P", `8` = "NEU_P", `9` = "NEU",  `10` = "NEU", `11` = "NEU", `12` = "NEU")
data_sub$celltype2 <- Idents(data_sub)

#visualize split by cell type  
Idents(data_sub)<- data_sub$ident 
pdf("cerebralorganoid_celltype_umap.pdf")
DimPlot(data_sub, reduction = "umap", label = T, pt.size = 0.2, group.by = "celltype2",  shuffle = T, raster = F, repel = T, label.size = 6)+theme(text = element_text(face = "bold"))+theme(title = element_text(size = 0))
dev.off()

#visualize split by dataset 
pdf("cerebralorganoid_comp_umap.pdf")
DimPlot(data_sub, reduction = "umap", label = F, pt.size = 0.2, group.by = "orig.ident",  shuffle = T, raster = F)+
  scale_color_manual(labels = c("AstTau", "Cerebral Organoid"),values = c("red","blue"))+theme(title = element_text(size = 0), legend.position = "bottom")
dev.off()
#make gradient function
make_gradient <- function(deg = 45, n = 100, cols = blues9) {
  cols <- colorRampPalette(cols)(n + 1)
  rad <- deg / (180 / pi)
  mat <- matrix(
    data = rep(seq(0, 1, length.out = n) * cos(rad), n),
    byrow = TRUE,
    ncol = n
  ) +
  matrix(
    data = rep(seq(0, 1, length.out = n) * sin(rad), n),
    byrow = FALSE,
    ncol = n
  )
  mat <- mat - min(mat)
  mat <- mat / max(mat)
  mat <- 1 + mat * n
  mat <- matrix(data = cols[round(mat)], ncol = n)
  grid::rasterGrob(
    image = mat,
    width = unit(1, "npc"),
    height = unit(1, "npc"), 
    interpolate = TRUE
  )
}
#count number of cells from each sample in each cluster
Idents(data_sub) <-"seurat_clusters"
ident <- c("integration","stress_organoid")
ns<-length(ident)
nb<-rep(0,ns) 
clusters<-levels(data_sub@active.ident)
ncl<-length(clusters)
counts<-matrix(0,nrow=ncl,ncol=ns)
rownames(counts)<-clusters
colnames(counts)<-ident
for (s in ident) {
  for (cl in clusters) {
    counts[cl,s]<-sum(data_sub@meta.data$orig.ident==s & data_sub@active.ident==cl)
  }
}
write.table(counts,file=paste0("liger_cell_counts.txt"),col.names=NA,sep="\t",quote=FALSE)

#manual gradient of least mature to most mature clusters 
colors<- c("red","red","orange","yellow","yellow","red","yellow","white","red","orange","red","red","orange","purple","yellow","orange","purple")
pdf("cerebralorganoid_comp_umap_gradient.pdf")
DimPlot(data_sub, reduction = "umap", label = F, pt.size = 0.2, group.by = "Subtype",cols = colors, shuffle = F, raster = F)+theme(plot.title =  element_text(size = 0))+ theme(legend.position = "none")
dev.off()

#reorder seurat cluster rownames to plot cell counts from least mature to most mature 
counts2 <- as.data.frame(counts)
counts2$names <- rownames(counts2)
counts3 <- counts2 %>% arrange(factor(names, levels = c('0','4','3','7','2','11','5','6','8','12','10','1','9')))
counts3.melt <- melt(counts3, id.var = 'names')
# plot histogram 
level_order = c('0','4','3','7','2','11','5','6','8','12','10','1','9')
g <- make_gradient(
  deg = 0, n = 500, cols = brewer.pal(12, "YlOrRd")
)
g2 <- ggplot(counts3.melt, aes(x = factor(names, level = level_order), y = value))+
  annotation_custom(
    grob = g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
  )
pdf("liger_comp_cellcounts.pdf")
g2 + geom_bar(aes(fill=variable), colour="white",stat = "identity", width = 0.5 ) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6, size = 10), axis.title.x = element_text(size=20), axis.title.y = element_text(size=20), legend.position = "bottom") + 
  labs(title="", 
       subtitle="") +
  scale_fill_manual(labels = c("AstTau", "Cerebral Organoid"),name = "Data Set",  values = c("red","blue") ) +labs(y= "Cell Counts", x = "Cell Cluster")
dev.off()


#STACKED VLN MARKER FIG

#NEW FUNCTION

## remove the x-axis text and tick
## plot.margin to adjust the white space between each plot.
## ... pass any arguments to VlnPlot in Seurat
modify_vlnplot<- function(obj, 
                          feature, 
                          pt.size = 0.0,
                          plot.margin = unit(c(0, 0, 0, 0), "cm"),
                          ...) {
  p<- VlnPlot(obj, features = feature, pt.size = pt.size, ... )  + 
    xlab("") + ylab(feature) + ggtitle("") + 
    theme(legend.position = "none", 
          plot.title = element_text(size = 0),plot.caption = element_text(size = 0),
          axis.text.x = element_text(size=0), axis.title.x = element_text(size=0),
          axis.ticks.length.x = unit(0, "pt"),
          axis.ticks.x = element_blank(),
          axis.title.y = element_text(size = rel(1), angle = 0, face = "bold"), 
          axis.text.y = element_text(size = rel(1), face = "bold"),
          plot.margin = plot.margin ) 
  return(p)
}

## extract the max value of the y axis
extract_max<- function(p){
  ymax<- max(ggplot_build(p)$layout$panel_scales_y[[1]]$range$range)
  return(ceiling(ymax))
}

## main function
StackedVlnPlot<- function(obj, features,
                          pt.size = 0.0, 
                          plot.margin = unit(c(0, 0, 0, 0), "cm"),
                          ...) {
  
  plot_list<- purrr::map(features, function(x) modify_vlnplot(obj = obj,feature = x, ...))
  
  # Add back x-axis title to bottom plot. patchwork is going to support this?
  plot_list[[length(plot_list)]]<- plot_list[[length(plot_list)]] +
    theme(axis.text.x=element_text(size = 15, angle = 35, face= "bold"), axis.ticks.x = element_line())
  
  # change the y-axis tick to only max value 
  ymaxs<- purrr::map_dbl(plot_list, extract_max)
  plot_list<- purrr::map2(plot_list, ymaxs, function(x,y) x +
                            scale_y_continuous(breaks = c(y)) +
                            expand_limits(y = y))

  p<- patchwork::wrap_plots(plotlist = plot_list, ncol = 1) 
        
  return(p)
}

features <-c("TOP2A","NES","VIM","SLC1A3","PEA15","GAP43","STMN2","SNAP25")
level <- c(0,4,3,7,2,11,5,6,8,12,10,1,9)

data <- SetIdent(data, value ="seurat_clusters")
data$seurat_clusters <- factor(x = data$seurat_clusters, levels = level)

data <- SetIdent(data, value ="seurat_clusters")
plot <- StackedVlnPlot(obj = data, features = features)

pdf("liger_cell_stckvln_celltype.pdf", width= 5, height =5 )
print(plot)
dev.off()

```

ONTOLOGY VLN PLOTS
```{r}
#WK3 ONLY 
modulesets <- c("ZAMANIAN_A1_AST","ZAMANIAN_A2_AST","tnf","ribo","HSF1","DAA_FIG2","GRUBMAN","GWAS")
celltypes <- c("IN_NEU","EX_NEU","ASC")
#V ONLY 
fillcolor <- c(col_ctr_v_w3, col_otau_v_w3)
for (moduleset in modulesets){
modulename=moduleset
module= eval(as.symbol(modulename))
for (ct in celltypes){
    tryCatch({
theMean <- colMeans(x = as.matrix(combined[['RNA']]@data[module, ]), na.rm = TRUE)
combined@meta.data$geneset <- theMean
combined=SetIdent(combined,value = "celltype.condition.week.treat")
subsetseurat <- subset(combined,idents=c(paste0(ct,"_ctr_w3_v"),paste0(ct,"_otau_w3_v")))
subsetseurat$celltype.condition.week.treat <- factor(as.character(subsetseurat$celltype.condition.week.treat), levels = c(paste0(ct,"_ctr_w3_v"),paste0(ct,"_otau_w3_v")))
comp.pairs <- list(c(paste0(ct,"_ctr_w3_v"),paste0(ct,"_otau_w3_v")))
p <- VlnPlot(
    object = subsetseurat, 
    features = "geneset", 
    group.by = "celltype.condition.week.treat", 
    cols = fillcolor,
    pt.size = 0.01) 
lim <-ggplot_build(p)$layout$panel_params[[1]]$y.range[2]
fun_mean <- function(x){  y = mean(x,na.rm=T)
  return(data.frame(y=lim+0.05,label=round(y,3)))}
pdf(paste0(modulename,"_",ct,"_v_avg_vio.pdf"))
print( p+ 
    ylim(0,lim+0.15)+
    stat_compare_means(label = "p.signif", 
                       method = "t.test", hide.ns = TRUE,
                       comparisons = comp.pairs,
                        label.y = c(lim,lim,lim), bracket.size = 1, size=25) +
geom_boxplot(width = 0.2, color="white", outlier.shape = NA, lwd=2) +   
stat_summary(fun.data = fun_mean, geom="text", color="black",size=10 )+
    theme(legend.text=element_text(size = 10), 
     axis.title.x = element_text(size = 0),
     axis.title.y = element_text(size = 30),
     axis.text.x = element_text(size = 0),
     axis.text.y = element_text(size = 25))+
      ylab("Gene Expression")+
    labs(title = NULL) +NoLegend())
dev.off()
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}
#ALL
fillcolor <- c(col_ctr_v_w3, col_otau_v_w3,col_ctr_pu_w3, col_otau_pu_w3 )
for (moduleset in modulesets){
modulename=moduleset
module= eval(as.symbol(modulename))
for (ct in celltypes){
    tryCatch({
theMean <- colMeans(x = as.matrix(combined[['RNA']]@data[module, ]), na.rm = TRUE)
combined@meta.data$geneset <- theMean
combined=SetIdent(combined,value = "celltype.condition.week.treat")
subsetseurat <- subset(combined,idents=c(paste0(ct,"_ctr_w3_v"),paste0(ct,"_otau_w3_v"),paste0(ct,"_ctr_w3_pu"),paste0(ct,"_otau_w3_pu")))
subsetseurat$celltype.condition.week.treat <- factor(as.character(subsetseurat$celltype.condition.week.treat), levels = c(paste0(ct,"_ctr_w3_v"),paste0(ct,"_otau_w3_v"),paste0(ct,"_ctr_w3_pu"),paste0(ct,"_otau_w3_pu")))
comp.pairs <- list(c(paste0(ct,"_ctr_w3_v"),paste0(ct,"_otau_w3_v")),c(paste0(ct,"_ctr_w3_pu"),paste0(ct,"_otau_w3_pu")),c(paste0(ct,"_otau_w3_v"),paste0(ct,"_otau_w3_pu")))
p <- VlnPlot(
    object = subsetseurat,
    features = "geneset",
    group.by = "celltype.condition.week.treat",
    cols = fillcolor,
    pt.size = 0.01)
lim <-ggplot_build(p)$layout$panel_params[[1]]$y.range[2]
fun_mean <- function(x){  y = mean(x,na.rm=T)
  return(data.frame(y=lim+0.05,label=round(y,3)))}
pdf(paste0(modulename,"_",ct,"_all_avg_vio_v2.pdf"))
print( p+
      ylim(0,lim+(lim/2))+
    stat_compare_means(label = "p.signif",
                       method = "t.test", hide.ns = TRUE,
                       comparisons = comp.pairs,
                       bracket.size = 1, 
                       size=20)+
                       stat_compare_means(method = "anova", size=5)+
                       geom_boxplot(width = 0.2, color="white", outlier.shape = NA, lwd=2) +
stat_summary(fun.data = fun_mean, geom="text", color="black",size=8 )+
    theme(legend.text=element_text(size = 10),
     axis.title.x = element_text(size = 0),
     axis.title.y = element_text(size = 30),
     axis.text.x = element_text(size = 0),
     axis.text.y = element_text(size = 25))+
     ylab("Gene Expression")+
    labs(title = NULL) +NoLegend())
dev.off()
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}}
```

NEURONAL SUBCLUSTER MONOCLE 3 PSEUDOTIME
```{r}

#subset to neurons
Idents(combined)<-"celltype"
subset <- subset(combined, idents=c("EX_NEU"))
#subset to vehicle only
Idents(subset)<-"condition.treat.week"
subset <- subset(subset, idents=c( "ctr_v_w1", "otau_v_w1","ctr_v_w2", "otau_v_w2","ctr_v_w3", "otau_v_w3"))

subset_ct <- FindVariableFeatures(subset, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(subset_ct)
subset_ct <- ScaleData(subset_ct, features = all.genes)
subset_ct <- RunPCA(subset_ct, verbose = FALSE)
subset_ct <- RunUMAP(subset_ct, reduction = "pca", dims = 1:15)
subset_ct <- FindNeighbors(subset_ct, dims = 1:15, verbose = FALSE)
subset_ct <- FindClusters(subset_ct, verbose = FALSE)

#Save seurat object
save(subset_ct,file="subset_GLUNEU.RData")

#load seurat object
load(file = "subset_GLUNEU.RData")
CTS <-  "GLU_NEU"
  
#setup
#Condition
Idents(subset_ct) <- "condition.treat.week"
# Define an order of cluster identities
condition_levels <-  c("ctr_v_w1", "otau_v_w1","ctr_v_w2", "otau_v_w2","ctr_v_w3", "otau_v_w3")

# Relevel object@ident
subset_ct@meta.data$condition.treat.week <- factor(x = subset_ct@meta.data$condition.treat.week, levels = condition_levels)
#colors
cond_color_sub <- c(col_ctr_v_w1,col_otau_v_w1,col_ctr_v_w2,col_otau_v_w2,col_ctr_v_w3,col_otau_v_w3)
#cond_color_sub <- c(col_ctr_v_w3,col_otau_v_w3)
cond_color_sub2 <- c(col_otau_v_w1,col_otau_v_w2,col_otau_v_w3)

newlabels  <- c("7 DIV3D Control","7 DIV3D oTAU","14 DIV3D Control","14 DIV3D oTAU","21 DIV3D Control","21 DIV3D oTAU")
newlabelswk<- c("7 DIV3D","14 DIV3D","21 DIV3D")

#umaps
Idents(subset_ct)<-"seurat_clusters"
pdf(paste0("./subset/",CTS,"/",CTS,"_cluster_UMAP.pdf"))
print(DimPlot(subset_ct))
dev.off()
Idents(subset_ct)<-"condition.treat.week"
pdf(paste0("./subset/",CTS,"/",CTS,"_condition_UMAP.pdf"))
print(DimPlot(subset_ct, shuffle = T, cols = cond_color_sub))
dev.off()

#monocle3

cds <-readRDS(paste0( "./subset/",CTS,"/",CTS,"cds.RData"))

#convert seurat object to monocle object 
cds <- as.cell_data_set(subset_ct, assay = "RNA")
## Calculate size factors using built-in function in monocle3
cds <- estimate_size_factors(cds)
## Add gene names into CDS
cds@rowRanges@elementMetadata@listData[["gene_short_name"]] <- rownames(subset_ct[["RNA"]])

#PCA and UMAP 
cds <- preprocess_cds(cds, num_dim = 50)

cds <- reduce_dimension(cds)

pdf(paste0("./subset/",CTS,"/",CTS,"_celltype_m3.pdf"))
plot_cells(cds, color_cells_by="celltype",  trajectory_graph_color = "grey33",label_cell_groups=FALSE, group_label_size = 0, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F)+theme(text = element_text(face="bold",size=15))+ theme(legend.title = element_blank())
dev.off()

pdf(paste0("./subset/",CTS,"/",CTS,"_condtreatweek_m3.pdf"))
plot_cells(cds, color_cells_by="condition.treat.week", trajectory_graph_color = "grey33", label_cell_groups=FALSE, group_label_size = 0, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F)+theme(text = element_text(face="bold",size=15))+scale_color_manual( 'Condition',values = cond_color_sub, labels = newlabels)+ theme(legend.title = element_blank())
dev.off

pdf(paste0("./subset/",CTS,"/",CTS,"_week_m3.pdf"))
plot_cells(cds, color_cells_by="week", trajectory_graph_color = "grey33", label_cell_groups=FALSE, group_label_size = 0, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F)+theme(text = element_text(face="bold",size=15))+scale_color_manual(values = cond_color_sub2, labels = newlabelswk)+ theme(legend.title = element_blank())
dev.off()

#cluster 
cds = cluster_cells(cds, resolution=1e-4)

pdf(paste0("./subset/",CTS,"/",CTS,"_pseudocluster_m3.pdf"))
plot_cells(cds, label_cell_groups=T, group_label_size = 10, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F, trajectory_graph_color = "grey33")+theme(text = element_text(face="bold",size=15))
dev.off()

#trajectory 
cds <- learn_graph(cds, use_partition = F)


pdf(paste0("./subset/",CTS,"/",CTS,"_trajectory_m3.pdf"))
plot_cells(cds,
           color_cells_by = "condition.treat.week",
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE)
dev.off()

#pseudotime 

plot_cells(cds, color_cells_by = "partition")

cds <- order_cells(cds)

pdf(paste0("./subset/",CTS,"/",CTS,"_pseudotime_m3.pdf"))
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)+theme(text = element_text(face="bold",size=15))
dev.off()

saveRDS(cds, paste0( "./subset/",CTS,"/",CTS,"cds.RData"))

#pseudo deg

res <- graph_test(cds, neighbor_graph="principal_graph", cores=4)
ids <- row.names(subset(res, q_value < 0.05))

gene_module_df <- find_gene_modules(cds[ids,], resolution=c(10^seq(-6,-1)))
write_csv(gene_module_df, paste0("./subset/",CTS,"/",CTS,"_genemoduledf_pseuo.csv"))

gene_module_df <-read_csv(paste0("./subset/",CTS,"/",CTS,"_genemoduledf_pseuo.csv"))

cell_group_df <- tibble::tibble(cell=row.names(colData(cds)),
                                cell_group=cds@clusters$UMAP$clusters)

agg_mat <- aggregate_gene_expression(cds, gene_module_df, cell_group_df)
row.names(agg_mat) <- stringr::str_c("Module ", row.names(agg_mat))
pdf(paste0("./subset/",CTS,"/",CTS,"_pseudomoduleheatmap_m3.pdf"))
print(pheatmap::pheatmap(agg_mat,
                   scale="column", clustering_method="ward.D2"))
dev.off()


#cluster count barplot pseudo clusters 

#table data 
t <- table(cds@colData$condition.treat.week, cds@clusters$UMAP$clusters)
#convert to percentage of total
t <- prop.table(t,margin = 1)
t <- t*100

rownames(t) <- newlabels
#document
write_csv(as.data.frame(t), paste0("./subset/",CTS,"/",CTS,"_pseudocluster_counts.csv"))
# melt the data frame for plotting
data.m <- melt(t, id.vars='clusters')
#plot
plot<- ggplot(data.m, aes(y = Var1, x = value, fill = as.character(Var2))) +geom_bar(stat = 'identity', width = 1,colour="black") +theme(text = element_text(face = "bold"),legend.text = element_text(size = 10, colour = "black"), axis.text.y = element_text(size = 10, colour = "black", angle = 0), axis.title.y = element_text(size = 0), axis.title.x = element_text(size = 0),legend.position = "bottom", legend.direction = "horizontal")+ theme(panel.background = element_blank())+ 
  labs(fill='Cluster')+ theme(aspect.ratio = .2)+ guides(fill = guide_legend(nrow = 1))

pdf(paste0("./subset/",CTS,"/",CTS,"_pseudocluster_barplot.pdf"), width = 6)
print(plot)
dev.off()

#plot cells 

gene_module_df <- read.csv(paste0("./subset/",CTS,"/",CTS,"_genemoduledf_pseuo.csv"))


pdf(paste0("./subset/",CTS,"/",CTS,"_mod12_m3.pdf"))
plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(12)),
           label_cell_groups=FALSE, show_trajectory_graph=T, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F, trajectory_graph_color = "grey33")+theme(text = element_text(face="bold",size=15))
dev.off()


#gpro

#12
x <- c("ACOT7","PARK7","ENO1","SLC25A33","HMGN2","NUDC","ATP5IF1","CCDC28B","YARS","SMIM12","TRAPPC3","LSM10","UTP11","EBNA1BP2","AKR1A1","PGM1","GSTM3","POLR3GL","JTB","DPM3","SCAMP3","NAXE","MRPL24","PFDN2","B4GALT3","NDUFS2","UCK2","CCDC181","MRPL55","EIPR1","DDX1","AC016722.3","SNRNP27","DGUOK","PCGF1","AUP1","SMIM37","CCDC115","IMP4","MZT2A","METTL5","OLA1","NUP35","HSPE1","METTL21A","ATIC","MRPL44","EXOSC7","MAP4","RRP9","TNNC1","ATG3","GAP43","RAB7A","MRPL3","SRPRB","DNAJC19","LINC02043","TCTEX1D2","MRFAP1","LINC02265","MRPL1","MRPS18C","PPA2","NDUFC1","LSM6","NDUFAF2","TBCA","HSPA9","MRPL22","RPL26L1","NOP16","NHP2","EEF1E1","C6orf52","MCUR1","HLA-A","MRPS18B","C6orf226","MRPL14","NDUFAF4","SNX3","PKIB","HDDC2","MPC1","AIMP2","GGCT","LSM5","PSMA2","MRPL32","TBRG4","TMEM120A","MDH2","SRI","BUD31","COPS6","POP7","ARF5","AKR1B1","RHEB","SMS","PRDX4","APOO","NBDY","TCEAL8","TSC22D3","IDH3G","EMD","DKC1","ELOC","MRPS28","POLR2K","TSTA3","EXOSC4","CYC1","BOP1","DMAC1","NDUFB6","NUDT2","CCDC107","NANS","RGS3","NDUFA8","TSSC4","CCDC34","COMMD9","FEN1","ARL2","CCDC85B","DRAP1","NDUFV1","NDUFS8","MRPL48","NDUFC2","CWC15","ZPR1","THYN1","CISD1","FAM241B","PPIF","DPCD","FBXL15","PRDX3","UROS","MLF2","AC079601.1","ARF3","C12orf10","AC084033.3","NDUFA12","ERP29","MRPL57","POLR1D","SERP2","NUDT15","SPRYD7","PSMB5","SPTSSA","NDUFB1","IFI27L1","APOPT1","IMP3","ZFAND6","RPS17","POLR3K","SNRNP25","METTL26","STUB1","FAHD1","NTHL1","FLYWCH2","NMRAL1","UBALD1","DEXI","RSL1D1","C16orf45","ALDOA","VKORC1","CFAP20","ATP6V0D1","NOB1","CFDP1","KARS","AC009119.2","TIMM22","YWHAE","C1QBP","PEMT","SLC25A39","ATP5MC1","NME1","SAP30BP","ARL16","MRPL12","PYCR1","MYL12B","IDH3B","DSTN","SMIM26","DUSP15","COMMD7","DYNLRB1","PKIG","PFDN4","TCEA2","ATP5F1D","C19orf24","MYDGF","C19orf70","TUBB4A","CD320","MRPL4","TMEM205","ECSIT","GADD45GIP1","MRPL34","REX1BD","BORCS8","NDUFA13","UQCRFS1","ECH1","PSMC4","BLVRB","TOMM40","AP2S1","TMEM160","RUVBL2","TSEN34","HSPBP1","ISOC2","CCDC106","ZSCAN1","HDHD5","RANBP1","DGCR6L","MIF","SNU13","SOD1","PSMG1","SUMO3")

#GPRO
gostres <- gost(query = x, 
                organism = "gp__zSEF_sD9Q_d1M", ordered_query = TRUE, 
                multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = T, 
                user_threshold = 0.05, correction_method = "fdr", 
                numeric_ns = "", as_short_link = FALSE)
gostres2 <- gostres$result 
gostres3 <- gostres2[1:13]
#write.csv 
write_csv(gostres3, paste0("./subset/",CTS,"/",CTS,"_mod12_m3_gpro.csv"))

#select desired databases
toMatch <- c("HALLMARK", "KEGG", "GO","REACTOME")
gostres$result=gostres$result[grepl(paste(toMatch, collapse = "|"),gostres$result$term_id),]

# modify the g:Profiler data frame
gp_mod = gostres$result[,c("query", "source", "term_id",                            
                           "term_name", "p_value", "query_size",                 
                           "intersection_size", "term_size",                     
                           "effective_domain_size", "intersection")]

gp_mod$GeneRatio = paste0(gp_mod$intersection_size,  "/", gp_mod$query_size)

gp_mod$BgRatio = paste0(gp_mod$term_size, "/", gp_mod$effective_domain_size)

names(gp_mod) = c("Cluster", "Category", "Description", "ID", "p.adjust",                
                  "query_size", "Count", "term_size", "effective_domain_size",           
                  "geneID", "GeneRatio", "BgRatio")

gp_mod$geneID = gsub(",", "/", gp_mod$geneID)
gp_mod$Cluster = gsub("query_1", "Module 12", gp_mod$Cluster)
row.names(gp_mod) = gp_mod$ID

# define as compareClusterResult object
gp_mod_cluster = new("compareClusterResult", compareClusterResult = gp_mod)

# define as enrichResult object
gp_mod_enrich  = new("enrichResult", result = gp_mod)

pdf(paste0("./subset/",CTS,"/",CTS,"_mod12_m3_gpro_heatmap.pdf"), width = 25, height = 8)
barplot(gp_mod_enrich, showCategory = 40, font.size = 15) + 
  ggplot2::facet_grid(~Cluster) +
  ggplot2::ylab("Intersection size")+theme(text = element_text(face="bold",size=15))
dev.off()

```

ASTRO SUBCLUSTER MONOCLE 3 PSEUDOTIME
```{r}

#subset to asc
Idents(combined)<-"celltype"
subset <- subset(combined, idents=c("ASC"))
#subset to vehicle only
Idents(subset)<-"condition.treat.week"
subset <- subset(subset, idents=c( "ctr_v_w1", "otau_v_w1","ctr_v_w2", "otau_v_w2","ctr_v_w3", "otau_v_w3"))

subset_ct <- FindVariableFeatures(subset, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(subset_ct)
subset_ct <- ScaleData(subset_ct, features = all.genes)
subset_ct <- RunPCA(subset_ct, verbose = FALSE)
subset_ct <- RunUMAP(subset_ct, reduction = "pca", dims = 1:15)
subset_ct <- FindNeighbors(subset_ct, dims = 1:15, verbose = FALSE)
subset_ct <- FindClusters(subset_ct, verbose = FALSE)

#Save seurat object
save(subset_ct,file="subset_ASCNEU.RData")

#load seurat object
load(file = "subset_ASCNEU.RData")
CTS <-  "ASC"
  
#setup
#Condition
Idents(subset_ct) <- "condition.treat.week"
# Define an order of cluster identities
condition_levels <-  c("ctr_v_w1", "otau_v_w1","ctr_v_w2", "otau_v_w2","ctr_v_w3", "otau_v_w3")

# Relevel object@ident
subset_ct@meta.data$condition.treat.week <- factor(x = subset_ct@meta.data$condition.treat.week, levels = condition_levels)
#colors
cond_color_sub <- c(col_ctr_v_w1,col_otau_v_w1,col_ctr_v_w2,col_otau_v_w2,col_ctr_v_w3,col_otau_v_w3)
cond_color_sub2 <- c(col_otau_v_w1,col_otau_v_w2,col_otau_v_w3)

newlabels  <- c("7 DIV3D Control","7 DIV3D oTAU","14 DIV3D Control","14 DIV3D oTAU","21 DIV3D Control","21 DIV3D oTAU")
newlabelswk<- c("7 DIV3D","14 DIV3D","21 DIV3D")

#umaps
Idents(subset_ct)<-"seurat_clusters"
pdf(paste0("./subset/",CTS,"/",CTS,"_cluster_UMAP.pdf"))
print(DimPlot(subset_ct))
dev.off()
Idents(subset_ct)<-"condition.treat.week"
pdf(paste0("./subset/",CTS,"/",CTS,"_condition_UMAP.pdf"))
print(DimPlot(subset_ct, shuffle = T, cols = cond_color_sub))
dev.off()

#monocle3
CTS <-  "ASC"
cds <-readRDS(paste0( "./subset/",CTS,"/",CTS,"cds.RData"))

#convert seurat object to monocle object 
cds <- as.cell_data_set(subset_ct, assay = "RNA")
## Calculate size factors using built-in function in monocle3
cds <- estimate_size_factors(cds)
## Add gene names into CDS
cds@rowRanges@elementMetadata@listData[["gene_short_name"]] <- rownames(subset_ct[["RNA"]])

#PCA and UMAP 
cds <- preprocess_cds(cds, num_dim = 50)
cds <- reduce_dimension(cds)

pdf(paste0("./subset/",CTS,"/",CTS,"_celltype_m3.pdf"))
plot_cells(cds, color_cells_by="celltype",  trajectory_graph_color = "grey33",label_cell_groups=FALSE, group_label_size = 0, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F)+theme(text = element_text(face="bold",size=15))+ theme(legend.title = element_blank())
dev.off()

pdf(paste0("./subset/",CTS,"/",CTS,"_condtreatweek_m3.pdf"))
plot_cells(cds, color_cells_by="condition.treat.week", trajectory_graph_color = "grey33", label_cell_groups=FALSE, group_label_size = 0, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F)+theme(text = element_text(face="bold",size=15))+scale_color_manual( 'Condition',values = cond_color_sub, labels = newlabels)+ theme(legend.title = element_blank())
dev.off

pdf(paste0("./subset/",CTS,"/",CTS,"_week_m3.pdf"))
plot_cells(cds, color_cells_by="week", trajectory_graph_color = "grey33", label_cell_groups=FALSE, group_label_size = 0, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F)+theme(text = element_text(face="bold",size=15))+scale_color_manual(values = cond_color_sub2, labels = newlabelswk)+ theme(legend.title = element_blank())
dev.off()

#cluster 
cds = cluster_cells(cds, resolution=1e-4)

pdf(paste0("./subset/",CTS,"/",CTS,"_pseudocluster_m3.pdf"))
plot_cells(cds, label_cell_groups=T, group_label_size = 10, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F, trajectory_graph_color = "grey33")+theme(text = element_text(face="bold",size=15))
dev.off()

#trajectory 
cds <- learn_graph(cds, use_partition = F)


pdf(paste0("./subset/",CTS,"/",CTS,"_trajectory_m3.pdf"))
plot_cells(cds,
           color_cells_by = "condition.treat.week",
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE)
dev.off()

#pseudotime 

plot_cells(cds, color_cells_by = "partition")

cds <- order_cells(cds)

pdf(paste0("./subset/",CTS,"/",CTS,"_pseudotime_m3.pdf"))
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)+theme(text = element_text(face="bold",size=15))
dev.off()

saveRDS(cds, paste0( "./subset/",CTS,"/",CTS,"cds.RData"))

#pseudo deg

res <- graph_test(cds, neighbor_graph="principal_graph", cores=4)
ids <- row.names(subset(res, q_value < 0.05))

gene_module_df <- find_gene_modules(cds[ids,], resolution=c(10^seq(-6,-1)))
write_csv(gene_module_df, paste0("./subset/",CTS,"/",CTS,"_genemoduledf_pseuo.csv"))

gene_module_df <-read_csv(paste0("./subset/",CTS,"/",CTS,"_genemoduledf_pseuo.csv"))

cell_group_df <- tibble::tibble(cell=row.names(colData(cds)),
                                cell_group=cds@clusters$UMAP$clusters)



agg_mat <- aggregate_gene_expression(cds, gene_module_df, cell_group_df)
row.names(agg_mat) <- stringr::str_c("Module ", row.names(agg_mat))

pdf(paste0("./subset/",CTS,"/",CTS,"_pseudomoduleheatmap_m3.pdf"))
print(pheatmap::pheatmap(agg_mat,
                   scale="column", clustering_method="ward.D2"))
dev.off()


#cluster count barplot pseudo clusters 

t <- table(cds@colData$condition.treat.week, cds@clusters$UMAP$clusters)
#convert to percentage of total
t <- prop.table(t,margin = 1)
t <- t*100

rownames(t) <- newlabels
#document
write_csv(as.data.frame(t), paste0("./subset/",CTS,"/",CTS,"_pseudocluster_counts.csv"))
# melt the data frame for plotting
data.m <- melt(t, id.vars='clusters')
#plot
plot<- ggplot(data.m, aes(y = Var1, x = value, fill = as.character(Var2))) +geom_bar(stat = 'identity', width = 1,colour="black") +theme(text = element_text(face = "bold"),legend.text = element_text(size = 10, colour = "black"), axis.text.y = element_text(size = 10, colour = "black", angle = 0), axis.title.y = element_text(size = 0), axis.title.x = element_text(size = 0),legend.position = "bottom", legend.direction = "horizontal")+ theme(panel.background = element_blank())+ 
  labs(fill='Cluster')+ theme(aspect.ratio = .2)+ guides(fill = guide_legend(nrow = 1))

pdf(paste0("./subset/",CTS,"/",CTS,"_pseudocluster_barplot.pdf"), width = 6)
print(plot)
dev.off()


#module analysis expanded 

#load gene module dataframe 
CTS <-  "ASC"
cds <-readRDS(paste0( "./subset/",CTS,"/",CTS,"cds.RData"))

gene_module_df <- read.csv(paste0("./subset/",CTS,"/",CTS,"_genemoduledf_pseuo.csv"))

#one modules defining one pseudo cluster manually selected from module heatmap
modules <- c(62,15,3,64,9,16,19,4,35,42,7,6,13,66,55)

#module feature plot 
for (mod in modules) {
pdf(paste0("./subset/",CTS,"/",CTS,"_mod_",mod,"_m3.pdf"))
print(plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(mod)),
           label_cell_groups=FALSE, show_trajectory_graph=T, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F, trajectory_graph_color = "grey33")+theme(text = element_text(face="bold",size=15)))
dev.off()
}


#gpro 

for (mod in modules) {
  
#get gene module list   
df2 <- gene_module_df[gene_module_df[, 2]==mod,]
x<- paste(df2$id)

#run gpro
gostres <- gost(query = x,
                organism = "gp__zSEF_sD9Q_d1M", ordered_query = TRUE,
                multi_query = FALSE, significant = T, exclude_iea = FALSE,
                measure_underrepresentation = FALSE, evcodes = T,
                user_threshold = 0.05, correction_method = "fdr",
                numeric_ns = "", as_short_link = FALSE)
gostres2 <- gostres$result
gostres3 <- gostres2[1:13]

#write.csv
write_csv(gostres3, paste0("./subset/",CTS,"/",CTS,"_mod_",mod,"_m3_gpro.csv"))

#select desired databases
toMatch <- c("HALLMARK", "KEGG", "GO","REACTOME")
gostres$result=gostres$result[grepl(paste(toMatch, collapse = "|"),gostres$result$term_id),]

# modify the g:Profiler data frame
gp_mod = gostres$result[,c("query", "source", "term_id",
                           "term_name", "p_value", "query_size",
                           "intersection_size", "term_size",
                           "effective_domain_size", "intersection")]

gp_mod$GeneRatio = paste0(gp_mod$intersection_size,  "/", gp_mod$query_size)

gp_mod$BgRatio = paste0(gp_mod$term_size, "/", gp_mod$effective_domain_size)

names(gp_mod) = c("Cluster", "Category", "Description", "ID", "p.adjust",
                  "query_size", "Count", "term_size", "effective_domain_size",
                  "geneID", "GeneRatio", "BgRatio")

gp_mod$geneID = gsub(",", "/", gp_mod$geneID)
gp_mod$Cluster = gsub("query_1", paste0("Module_",mod), gp_mod$Cluster)
row.names(gp_mod) = gp_mod$ID

# define as compareClusterResult object
gp_mod_cluster = new("compareClusterResult", compareClusterResult = gp_mod)

# define as enrichResult object
gp_mod_enrich  = new("enrichResult", result = gp_mod)

pdf(paste0("./subset/",CTS,"/",CTS,"_mod_",mod,"_m3_gpro_heatmap.pdf"), width = 15, height = 8)
print(barplot(gp_mod_enrich, showCategory = 40, font.size = 15) +
  ggplot2::facet_grid(~Cluster) +
  ggplot2::ylab("Intersection size")+theme(text = element_text(face="bold",size=15)))
dev.off()

}

#42
x <- c("APOE","SLC6A11","CTSH","LINC00882","MSTN","KCNJ16","MMP14","SLC25A48","LINC01550","CTXND1","EYA2","RAB27B","FDX2","DUSP23","KLF1","NPY","NUPR1","CENPVL3","MSRA","EPHB6","ALDOC","LINC00645","C9orf106","MFGE8","SGCD","CST3","CCL2","PFKP","AL133351.1","LINC01579","RARRES3","CCDC13","SHISA9","SRPX","RAMP1","FIBIN","MMP28","PTGDS","APOL4","LINC00844","NAGS","PTCHD1","C16orf74","CPNE5","BOLA1","NOG","MMD2","EFEMP1","CMTM5","GAS2L1","PLEKHB1","ASTN2","GAS7","DNAH9","HEPN1","GRIA3","PRDM16","NDUFA7","BRINP1","GPER1","SERPINI2","HEPACAM","RAP1GAP","EFEMP2","TRIM47","AL122058.1","CDH13","AL590617.2","AC005476.2","SLC35F4","ANKFN1","RBFOX1","TLCD1","SEMA6D","BBOX1","ERICH2","IQSEC2","DBX2","KANK1","MAOB","MRO","MFAP4","MT3","ACADL","SCN4B","HSPB1","TMEM100","NDRG2","SLC22A18","EMC10","TMEM63B","HSPB8","AC006960.3","RASSF10","TFPI","EYA4","GMPR","BST2","KCNIP2","ZNF843","HNMT","CD38","S100A16","AP001626.1","RWDD3","TSC22D4","TAP1","DBI","SNCG","ASTN1","PMP2","RABIF","GSTM4","SLITRK2","GABBR2","SLC7A10","CAPS2","AL451165.2","SLC6A1","ME1","ARHGAP19-SLIT1","ADD3","MB","MEGF11","PDLIM4","STOM","EGFR","GNG7","PSMB10","HDHD3","BRINP3","AC025171.3","NFIA","AIFM3","VSTM4","KCNIP1","ACSL6","AC007325.4","TIMP4","DTNA","NFIC","CAMK2G","LINC01094","SCARB1","CCDC88B","ELN","PQLC3","BHLHE41","AC007423.1","MCRIP2","UBTD1","CARD19","DTX3L")

pdf(paste0("./subset/",CTS,"/",CTS,"_mod_15_16_m3.pdf"), width = 10)
print(plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(15,16)),
           label_cell_groups=FALSE, show_trajectory_graph=F, label_principal_points = F, label_leaves = F, label_roots = F, label_branch_points = F, trajectory_graph_color = "grey33")+theme(text = element_text(face="bold",size=15)))
dev.off()

```

EPHYS GENES
```{r}

#paths <- c("vg_ca_c","chol_r","cl_ch","dop_r","gaba_r","glyc_r","ion_glu_r","meta_glu_r","cycl_nucleo_c","ser_r","vg_k_c","rect_k_c","twin_k_c","ca_act_k_c","ryan_r","vg_na_c")

#paths <- c("vg_ca_c","cl_ch","cycl_nucleo_c","vg_k_c","rect_k_c","twin_k_c","ca_act_k_c","vg_na_c")

#concatenate all channel paths 
newpath <- append(vg_ca_c,cl_ch)
newpath <- append(newpath, cycl_nucleo_c)
newpath <- append(newpath, vg_k_c)
newpath <- append(newpath,rect_k_c)
newpath <- append(newpath,twin_k_c)
newpath <- append(newpath,ca_act_k_c)
newpath <- append(newpath,vg_na_c)

#for plotting all

genes <- newpath
  
temp <- combined

DefaultAssay(temp) <- "RNA"

genes = genes[genes %in% rownames(temp@assays$RNA@data)]

temp=SetIdent(temp,value="celltype")

heatcolor=brewer.pal(11,"RdYlBu")

what=AverageExpression(temp,features=genes,verbose=F,slot="scale.data")$RNA

min=quantile(unlist(what),0.1)
mid=quantile(unlist(what),0.5)
max=quantile(unlist(what),0.9)

what <- what[,c(4,5,2)]

p <- Heatmap(what,
             width = unit(6, "cm"),
             height = unit((.35*length(genes)),"cm"),
             name="Expression",
             col=colorRamp2(c(min, 0, max), c("blue", "white", "red")),
             clustering_method_columns = "ward.D2",
             cluster_rows=FALSE,
             cluster_columns=FALSE,
             show_column_names=T,
             show_row_names=T,
             rect_gp = gpar(col = "black", lwd = 2), 
             show_column_dend = FALSE, 
             show_row_dend = FALSE, 
             column_names_rot = 0, 
             column_names_centered = T,
             column_title_gp = gpar(fontface="bold"), 
             column_names_gp = gpar(fontface="bold"), 
             row_names_gp = gpar(fontface="bold"))

ht_list = p
p <-draw(p, column_title = "Ion Channel Expression",column_title_gp = gpar(fontface = "bold"))

pdf(paste0("./ephys/all_heatmap.pdf"), height = (.5*length(genes)))
print(p)
dev.off()



#for plotting split by type

for (path in paths) {
  
modulename=path
genes= eval(as.symbol(modulename))
  
temp <- combined

DefaultAssay(temp) <- "RNA"

genes = genes[genes %in% rownames(temp@assays$RNA@data)]

temp=SetIdent(temp,value="celltype")

heatcolor=brewer.pal(11,"RdYlBu")
what=AverageExpression(temp,features=genes,verbose=F,slot="scale.data")$RNA
min=quantile(unlist(what),0.1)
mid=quantile(unlist(what),0.5)
max=quantile(unlist(what),0.9)

what <- what[,c(1,3,2)]

p <- Heatmap(what,
             width = unit(6, "cm"),
             height = unit((.5*length(genes)),"cm"),
             name=modulename,
             col=colorRamp2(c(min, 0, max), c("blue", "white", "red")),
             clustering_method_columns = "ward.D2",
             cluster_rows=FALSE,
             cluster_columns=FALSE,
             show_column_names=T,
             show_row_names=T,
             rect_gp = gpar(col = "black", lwd = 2), 
             show_column_dend = FALSE, 
             show_row_dend = FALSE, 
             column_names_rot = 0, 
             column_names_centered = T,
             column_title_gp = gpar(fontface="bold"), 
             column_names_gp = gpar(fontface="bold"), 
             row_names_gp = gpar(fontface="bold"))

ht_list = p
p <-draw(p, column_title = modulename,column_title_gp = gpar(fontface = "bold"))

pdf(paste0("./ephys/",modulename,"_heatmap.pdf"), height = (1*length(genes)))
print(p)
dev.off()

}

#genes

#voltage gated calcium 
vg_ca_c <- c("CACNA1A","CACNA1B","CACNA1C","CACNA1D","CACNA1E","CACNA1F","CACNA1G","CACNA1H","CACNA1I","CACNA1S","CACNA2D1","CACNA2D2","CACNA2D3","CACNA2D4","CACNB1","CACNB2","CACNB3","CACNB4","CACNG1","CACNG2","CACNG3","CACNG4","CACNG5","CACNG6","CACNG7","CACNG8")
#subset <- AddModuleScore(object=subset, features = vg_ca_c, ctrl = 5, name = "vg_ca_c")

#cholinergic receptors 
chol_r <- c("CHRNA1","CHRNA10","CHRNA2","CHRNA3","CHRNA4","CHRNA5","CHRNA6","CHRNA7","CHRNA9","CHRNB1","CHRNB2","CHRNB4","CHRND","CHRNE","CHRNG")
#"CHRNB3"
#subset <- AddModuleScore(object=subset, features = chol_r, ctrl = 5, name = "chol_r")

#chloride channels 
cl_ch <- c("CLCN1","CLCN2","CLCN3","CLCN4","CLCN5","CLCN6","CLCN7","CLCNKB")
#"CLCNKA",
#subset <- AddModuleScore(object=subset, features = cl_ch, ctrl = 5, name = "cl_ch")

#dopamine receptors 
dop_r <- c("DRD2","DRD4","DRD5")
#"DRD1","DRD3"
#subset <- AddModuleScore(object=subset, features = dop_r, ctrl = 5, name = "dop_r")

#gaba receptors
gaba_r <- c("GABBR1","GABBR2","GABRA1","GABRA2","GABRA3","GABRA4","GABRA5","GABRA6","GABRB1","GABRB2","GABRB3","GABRD","GABRE","GABRG1","GABRG2","GABRG3","GABRP","GABRQ","GABRR1","GABRR2")
#subset <- AddModuleScore(object=subset, features = gaba_r, ctrl = 5, name = "gaba_r")

#glycine receptors 
glyc_r <- c("GLRA1","GLRA2","GLRA3","GLRB","GRIA1","GRIA2","GRIA3","GRIA4")
#"GCOM1",
#subset <- AddModuleScore(object=subset, features = glyc_r, ctrl = 5, name = "glyc_r")

#ionotropic glutamate receptors 
ion_glu_r <- c("GRID1","GRID2","GRIK1","GRIK2","GRIK3","GRIK4","GRIK5","GRIN1","GRIN2A","GRIN2B","GRIN2C","GRIN2D","GRIN3A","GRINA")
#subset <- AddModuleScore(object=subset, features = ion_glu_r, ctrl = 5, name = "ion_glu_r")

#metabotropic glutatmate receptors
meta_glu_r <- c("GRM1","GRM2","GRM3","GRM4","GRM5","GRM6","GRM7","GRM8")
#subset <- AddModuleScore(object=subset, features = meta_glu_r, ctrl = 5, name = "meta_glu_r")

#cyclic nucleotide gated channel 
cycl_nucleo_c <- c("HCN1","HCN2","HCN3","HCN4")
#subset <- AddModuleScore(object=subset, features = cycl_nucleo_c, ctrl = 5, name = "cycl_nucleo_c")

#serotonin receptors
ser_r <- c("HTR1B","HTR1D","HTR1E","HTR1F","HTR2A","HTR2C","HTR3B","HTR3E","HTR7")
#"HTR1A","HTR3A","HTR3C","HTR3D","HTR4","HTR5A","HTR6"
#subset <- AddModuleScore(object=subset, features = ser_r, ctrl = 5, name = "ser_r")

#voltage gated potassium channels 
vg_k_c <-c("KCNA1","KCNA2","KCNA3","KCNA4","KCNA5","KCNA7","KCNAB1","KCNAB2","KCNAB3","KCNB1","KCNB2","KCNC1","KCNC2","KCNC3","KCNC4","KCND1","KCND2","KCND3","KCNE1","KCNE2","KCNE3","KCNE4","KCNF1","KCNG1","KCNG2","KCNG3","KCNG4","KCNH1","KCNH2","KCNH3","KCNH4","KCNH5","KCNH6","KCNH7","KCNH8","KCNQ1","KCNQ2","KCNQ3","KCNQ4","KCNQ5","KCNRG","KCNS1","KCNS2","KCNS3","KCNT1","KCNV1","KCNV2")
#"KCNA10","KCNA6","KCNE1L",
#subset <- AddModuleScore(object=subset, features = vg_k_c, ctrl = 5, name = "vg_k_c")

#potassium inward rectifying channel 
rect_k_c <- c("KCNJ1","KCNJ10","KCNJ11","KCNJ12","KCNJ14","KCNJ15","KCNJ16","KCNJ2","KCNJ3","KCNJ4","KCNJ5","KCNJ6","KCNJ8","KCNJ9")
#subset <- AddModuleScore(object=subset, features = rect_k_c, ctrl = 5, name = "rect_k_c")

#twin pore potassium channel
twin_k_c <- c("KCNK1","KCNK10","KCNK12","KCNK13","KCNK15","KCNK17","KCNK2","KCNK3","KCNK4","KCNK5","KCNK6","KCNK7","KCNK9")
#subset <- AddModuleScore(object=subset, features = twin_k_c, ctrl = 5, name = "twin_k_c")
#"KCNK16"

#calcium activated potassium channel
ca_act_k_c <- c("KCNMA1","KCNMB2","KCNMB3","KCNMB4","KCNN1","KCNN2","KCNN3","KCNN4")
#subset <- AddModuleScore(object=subset, features = ca_act_k_c, ctrl = 5, name = "ca_act_k_c")

#ryanodine receptor
ryan_r <- c("RYR1","RYR2","RYR3")
#subset <- AddModuleScore(object=subset, features = ryan_r, ctrl = 5, name = "ryan_r")

#voltage gated sodium channel genes 
vg_na_c <- c("SCN11A","SCN1B","SCN2B","SCN3A","SCN3B","SCN4B","SCN5A","SCN7A","SCN8A","SCN9A","SCN1A")
#"SCN10A","SCN2A2","SCN4A"
#subset <- AddModuleScore(object=subset, features = vg_na_c, ctrl = 5, name = "vg_na_c")



#plot


#paths <- c("vg_ca_c","chol_r","cl_ch","dop_r","gaba_r","glyc_r","ion_glu_r","meta_glu_r","cycl_nucleo_c","ser_r","vg_k_c","rect_k_c","twin_k_c","ca_act_k_c","ryan_r","vg_na_c")

paths <- c("vg_ca_c","cl_ch","cycl_nucleo_c","vg_k_c","rect_k_c","twin_k_c","ca_act_k_c","vg_na_c")

#pathway <- c(vg_ca_c,chol_r,cl_ch,dop_r,gaba_r,glyc_r,ion_glu_r,meta_glu_r,cycl_nucleo_c,ser_r,vg_k_c,rect_k_c,twin_k_c,ca_act_k_c,ryan_r,vg_na_c)


#feature plot 
for (path in paths) {
  
pdf(paste0("./ephys/",path,"_feature.pdf"))

print(FeaturePlot(subset, features = paste0(path,"1"), min.cutoff = 0, order = T, raster = F)+theme(text = element_text(face = "bold"))+labs(title = path))

dev.off()

}





```

SYNAPTIC MARKERS
```{r}

synaptic <- c("RAB14","DLG4","SNCA","SYN1","NRXN3","STX1A","NLGN1","HOMER1", "GAP43","SYP","SNAP25")

#dot plot 

conds <- c("ctr","otau")
combined=SetIdent(combined,value="condition")
subset=subset(combined,idents=conds)

synaptic <- c("SYN1","NRXN3","RAB14","DLG4","SNCA","STX1A","NLGN1","HOMER1")

plot<- DotPlot(subset, features = synaptic , cluster.idents = F, cols = c("blue", "red"), dot.scale = 15,scale.by = "size")+
  theme(text=element_text(face="bold"),axis.text.x = element_text(size=10, angle = 90), legend.position = "bottom", legend.text = element_text(size = 10), legend.title = element_text(size = 10))+  labs(y="Condition",x="Genes")+theme(aspect.ratio = .2)

pdf("./gwas/synaptic_dotplot.pdf")
print(plot)
dev.off()
```
                                                                                                                                          SEURAT OBJECT FILE OUTPUT 
```{r}
#for Broad Single Cell Portal Submission 
#metadata- rename CELL_1... in excel and add additional required columns 
write.csv(combined@meta.data, file = "asteroid1_meta.csv")
#reducation- rename CELL_1... in excel 
write.csv(combined@reductions$umap@cell.embeddings, file = "asteroid1_reductions.csv")
#normalized data matrix 
memory.size(max=TRUE)
a=combined@assays$RNA@data
colnames(a)=paste0("CELL_",1:ncol(a))
#count matrix - UMI corrected raw counts 
memory.size(max=TRUE)
a=combined@assays$RNA@counts
colnames(a)=paste0("CELL_",1:ncol(a))
write10xCounts(path="count_matrix", type = "sparse",x = a) 

#for NIH GEO Submission 
#count matrix - UMI corrected raw counts 
memory.size(max=TRUE)
a=combined@assays$RNA@counts
write10xCounts(path="count_matrix_GEO", type = "sparse",x = a) 
```